---
title: "Cody"
author: "Cody Spiegel"
date: "4/22/2022"
output: html_document
---

```{r setup chunk, setup, include = FALSE, cache=FALSE, message=FALSE, warning=FALSE}
if (!require('knitr')) install.packages('knitr'); library('knitr')
knitr::opts_chunk$set(warning=FALSE, message=FALSE, fig.align='center')

# load packages
if (!require("pacman")) install.packages("pacman") # for rapid install if not in library

# use pacman to load all the packages you are missing!
pacman::p_load('knitr', 'lme4', 'lmerTest', 'tidyverse', 'magrittr', 'effects', 'plyr', 'dplyr', 
               'plotrix', 'car',"gridExtra", "cowplot", "tools")
Fig.formatting<-(theme_classic()) +
  theme(text=element_text(size=10),
        legend.text.align = 0,
        panel.border = element_rect(fill=NA, colour = "black", size=1),
        aspect.ratio=1, 
        axis.ticks.length=unit(0.25, "cm"),
        axis.text.y=element_text(
          margin=unit(c(0.5, 0.5, 0.5, 0.5), "cm"), colour="black", size=10), 
        axis.text.x=element_text(
          margin=unit(c(0.5, 0.5, 0.5, 0.5), "cm"), colour="black", size=8)) +
  theme(legend.key.size = unit(0.4, "cm")) +
  theme(aspect.ratio=1.3) +
  theme(panel.spacing=unit(c(0, 0, 0, 0), "cm"))


```

###DOC  
import and do a loop to clean up all files and make stacked data in single df
```{r}
## import treatment IDs
IDs<-read.csv("/Users/tobyspiegel2/Desktop/My_pyromania/Pyromania/data/treatment.IDs.csv")

##### grab files in a list
Total.DOC.files <- list.files(path="/Users/tobyspiegel2/Desktop/My_pyromania/Pyromania/data/DOC.TN", pattern = "csv$", full.names = T)

##### what are the file names, sans extensions using package 'tools'
file.names<-file_path_sans_ext(list.files(path="/Users/tobyspiegel2/Desktop/My_pyromania/Pyromania/data/DOC.TN", pattern = "csv$", full.names = F))

############ formatting all data in for loop
  for(i in 1:length(Total.DOC.files))
    {
  data<-read.csv(Total.DOC.files[i], sep=",")
  data<-data[,c(1,3,4)] # removed columns we don't need
  data$File<-Total.DOC.files[i]
  colnames(data)<-c("Tank", "DOC..mg.L", "TN..mg.L", "File")
  data$Tank<- IDs$Tank
  data$Tank<-as.numeric(as.character(data$Tank)) # make the column of samples all numeric
  data <- data[!is.na(as.numeric(as.character(data$Tank))),] # remove all rows that aren't numeric/tanks
  data$Treatment<-IDs$Treatment
  data$plant.mass..g<-IDs$plant.mass..g
  make.names(assign(paste(file.names[i], sep=""), data)) # make the file name the name of new df for loop df
  }

########## this is the end of the loop

#Combine files from loop to single df
DOC.df<-rbind(DOC_T1, DOC_T2, DOC_T3, DOC_T4)
DOC.df$Treatment=as.factor(DOC.df$Treatment)

DOC.df$File<-substr(DOC.df$File, 64, 74) 
# alternative way to code the above
#give the 10th-24th character of the file name, removing the rest
#DOC.df$File<-substr(DOC.df$File, 13, 27) 

#alternatively
# remove the 9 letters ('^.) at start 
# remove the 4 letters (.$') at end
#DOC.df$File<-gsub('^.........|....$', '', DOC.df$File) 

# if equals DOC_T0_11052021 then, T0, if not then T1
DOC.df$Time.point<- as.factor(
         ifelse (DOC.df$File=="DOC_T1.csv", "T1",
           ifelse (DOC.df$File=="DOC_T2.csv", "T2", 
                   ifelse(DOC.df$File=="DOC_T3.csv", "T3", "T4"))))

#rearrange
library(dplyr)
DOC.df<- DOC.df %>% 
  dplyr::select(File, Time.point, Treatment, Tank, plant.mass..g, DOC..mg.L, TN..mg.L) 

#visualize
ggplot(DOC.df, aes(x = plant.mass..g, y = DOC..mg.L, colour = Treatment)) +
    geom_point() +
    facet_grid(~Time.point)+
    geom_smooth(method = 'loess', se = FALSE) +
    scale_colour_brewer(type = 'qual', palette = 'Dark2') +
    theme(legend.position = 'top')


```


```{r}
#A single common smoother plus group-level smoothers that have the same wiggliness (model GS)
m1.gam=gam(log(DOC..mg.L) ~ Treatment + s(plant.mass..g, k=8, m=1) + s(plant.mass..g, factor(Time.point), k=8, bs="fs", m=1), data=DOC.df, method="REML")
summary.gam(m1.gam)
draw(m1.gam)
k.check(m1.gam)
#all groups have similar functional responses, but intergroup variation in responses is allowed. This approach works by allowing each grouping level to have its own functional response, but penalizing functions that are too far from the average.

```

```{r}
#add in a smooth term. This is a thin plate regression spline, where we are smoothing over mass
m1.gam=gam(log(DOC..mg.L)~ s(log(plant.mass..g), m=2, k=8) + s(), method="REML", family= gaussian(link="log"), data = DOC_T1)
summary(m1.gam)
gam.check(m1.gam, rep=1000)
plot(m1.gam)

gam(log(uptake) ∼ s(log(conc), k=5, m=2) +
s(log(conc), Plant_uo, k=5, bs="fs", m=2),
data=CO2, method="REML")


m1.lm=lm(DOC..mg.L ~ Treatment+ poly(plant.mass..g,3), data=DOC_T1)
summary(m1.lm)
qqnorm(resid(m1.lm))
qqline(resid(m1.lm))




```{r}
#observations have some distribution
#wiggliness is smoothing curve
#assumptions for glm hold true for GAM...must be satisfied 

####Gamma distribution
#model 1: basic model fit to all data
m1.gam=gam(DOC..mg.L~  
             s(plant.mass..g) + #smooth of plant mass trend
             s(Time.point, bs="re"),#random effect of time point
           data= DOC.df,
           family= Gamma(link = "log"),#for positive skewed data
           methods= 'REML',
           select=TRUE)
summary(m1.gam)

```


Model 2
```{r}
#model 2: by treatment
m2.gam=gam(DOC..mg.L~ Treatment + #main effect
             s(plant.mass..g,  by=Treatment, k=8, m=1) + #treatment specific trends
             s(Time.point, bs="re", k=8),#random effect of time point 
           data= DOC.df,
           family= Gamma(link = "log"),#for positive skewed data
           methods= 'REML',
           select=TRUE)
summary(m2.gam)
draw(m2.gam)
```
Model 3
```{r}
#model 3: individual trends
m3.gam=gam(DOC..mg.L~ Treatment + 
             s(plant.mass..g, Time.point, bs="fs", k=8, m=1),#time point specific trends
           data= DOC.df,
           family= Gamma(link = "log"),#for positive skewed data
           methods= 'REML',
           select=TRUE)
summary(m3.gam)
draw(m3.gam)
```
Model 4
```{r}
#model 4: treatment specific trends, plus individual trends
m4.gam=gam(DOC..mg.L~  Treatment+ s(plant.mass..g, bs="tp", k=8)+ #main effect
             s(plant.mass..g, by= Treatment, k=8, m=1, bs="fs") + #treatment specific trends
             s(plant.mass..g, Time.point, bs="fs", k=8, m=1),#time point specific trends as random effect
           data= DOC.df,
           family= Gamma(link = "log"),#for positive skewed data
           methods= 'REML', 
           select=TRUE)#more penalization
summary(m4.gam)
draw(m4.gam)
gam.check(m4.gam, rep= 1000)##reps provide confidence intervals

?anova.gam

library(gratia)#gavin's package the uses ggplot as a base to see residuals
draw(m4.gam, residuals = T)#looking at smoothing effects
k.check(m4.gam)#checks to see if the k is good..and they are
```
Log transformed (model 4)
```{r}
#model 5 
m5.gam= gam(log10(DOC..mg.L)~ Treatment +
              s(plant.mass..g, by= Treatment, k=15)+ 
               s(plant.mass..g, Time.point, bs="fs", k=15),
           data= DOC.df,
           family= gaussian(),
           methods= 'REML',
           select= TRUE)
summary(m5.gam)
draw(m5.gam)
gam.check(m5.gam, rep=1000)#reps provide confidence intervals
