---
title: "Pyromania"
author: "C Wall"
date: "1/4/2021"
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 4
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '4'
---

```{r setup chunk, setup, include = FALSE, cache=FALSE, message=FALSE, warning=FALSE}
if (!require('knitr')) install.packages('knitr'); library('knitr')
knitr::opts_chunk$set(warning=FALSE, message=FALSE, fig.align='center')

# load packages
if (!require("pacman")) install.packages("pacman") # for rapid install if not in library

# use pacman to load all the packages you are missing!
pacman::p_load('knitr', 'lme4', 'lmerTest', 'tidyverse', 'magrittr', 'effects', 'plyr', 'dplyr', 
               'plotrix', 'car',"gridExtra", "cowplot", "tools")


```


###YSI  
Time 1 and Time 2: import YSI data and produce plots of changes in O2% and net ecosystem productivity (NEP) and net ecosystem respiration (NER)  
```{r YSI organized}
#load YSI data
YSI<-read.csv("data/Pyro_YSI.csv")

# fix date
YSI$Date<-as.character(YSI$Date)
YSI$Date<-as.POSIXct(YSI$Date, format="%m/%d/%Y")
YSI$Date<-as.Date(YSI$Date, format="%m/%d/%Y")


####### Time 1 change in O2 ################

#separate time points
YSI.T1<- YSI[(YSI$Time.point=="T1"),]

#calculate NEP for T1
T1.Prod<-YSI.T1[(YSI.T1$Date == "2021-11-15"),] # dawn and dusk for 12h period
T1.Dawn1<-T1.Prod[(T1.Prod$Dawn..Dusk == "dawn"),] # dawn-1 measurements
T1.Dusk<-T1.Prod[(T1.Prod$Dawn..Dusk == "dusk"),] # dusk measurements

T1.Dawn2<-YSI.T1[(YSI.T1$Date == "2021-11-16"),] # dawn-2 measurements, following AM

# make new dataframe
T1.O2<-(T1.Dawn1[,c(2,4:6)]) 
T1.O2$dawn1<-T1.Dawn1$DO.percent
T1.O2$dusk1<-T1.Dusk$DO.percent
T1.O2$dawn2<-T1.Dawn2$DO.percent

# NER = dusk - dawn (PM to AM, O2 change of day 1)
# NEP = dusk - dawn (PM to AM, O2 change of day 2)

T1.O2<- mutate(T1.O2, 
                NEP=dusk1 - dawn1,
                NER=dawn2 - dusk1) 

#sort
T1.O2<-T1.O2 %>% 
  arrange(Treatment, plant.mass..g) 


################ ################ ################
####### Time 2 change in O2 ################

#separate time points
YSI.T2<- YSI[(YSI$Time.point=="T2"),]

#calculate NEP for T2
T2.Prod<-YSI.T2[(YSI.T2$Date == "2021-12-06"),] # dawn and dusk for 12h period
T2.Dawn1<-T2.Prod[(T2.Prod$Dawn..Dusk == "dawn"),] # dawn-1 measurements
T2.Dusk<-T2.Prod[(T2.Prod$Dawn..Dusk == "dusk"),] # dusk measurements

T2.Dawn2<-YSI.T2[(YSI.T2$Date == "2021-12-07"),] # dawn-2 measurements, following AM

# make new dataframe
T2.O2<-(T2.Dawn1[,c(2,4:6)]) 
T2.O2$dawn1<-T2.Dawn1$DO.percent
T2.O2$dusk1<-T2.Dusk$DO.percent
T2.O2$dawn2<-T2.Dawn2$DO.percent

T2.O2<- mutate(T2.O2, 
                NEP=dusk1 - dawn1,
                NER=dawn2 - dusk1) 

#sort
T2.O2<-T2.O2 %>% 
  arrange(Treatment, plant.mass..g) 


################ ################ ################
####### Time 3 change in O2 ################

#separate time points
YSI.T3<- YSI[(YSI$Time.point=="T3"),]

#calculate NEP for T3
T3.Prod<-YSI.T3[(YSI.T3$Date == "2022-01-03"),] # dawn and dusk for 12h period
T3.Dawn1<-T3.Prod[(T3.Prod$Dawn..Dusk == "dawn"),] # dawn-1 measurements
T3.Dusk<-T3.Prod[(T3.Prod$Dawn..Dusk == "dusk"),] # dusk measurements

T3.Dawn2<-YSI.T3[(YSI.T3$Date == "2022-01-04"),] # dawn-2 measurements, following AM

# make new dataframe
T3.O2<-(T3.Dawn1[,c(2,4:6)]) 
T3.O2$dawn1<-T3.Dawn1$DO.percent
T3.O2$dusk1<-T3.Dusk$DO.percent
T3.O2$dawn2<-T3.Dawn2$DO.percent

T3.O2<- mutate(T3.O2, 
                NEP=dusk1 - dawn1,
                NER=dawn2 - dusk1) 

#sort
T3.O2<-T3.O2 %>% 
  arrange(Treatment, plant.mass..g) 

################ ################ ################
####### Time 3 change in O2 ################

#separate time points
YSI.T4<- YSI[(YSI$Time.point=="T4"),]

#calculate NEP for T4
T4.Prod<-YSI.T4[(YSI.T4$Date == "2022-02-02"),] # dawn and dusk for 12h period
T4.Dawn1<-T4.Prod[(T4.Prod$Dawn..Dusk == "dawn"),] # dawn-1 measurements
T4.Dusk<-T4.Prod[(T4.Prod$Dawn..Dusk == "dusk"),] # dusk measurements

T4.Dawn2<-YSI.T4[(YSI.T4$Date == "2022-02-03"),] # dawn-2 measurements, following AM

# make new dataframe
T4.O2<-(T4.Dawn1[,c(2,4:6)]) 
T4.O2$dawn1<-T4.Dawn1$DO.percent
T4.O2$dusk1<-T4.Dusk$DO.percent
T4.O2$dawn2<-T4.Dawn2$DO.percent

T4.O2<- mutate(T4.O2, 
                NEP=dusk1 - dawn1,
                NER=dawn2 - dusk1) 

#sort
T4.O2<-T4.O2 %>% 
  arrange(Treatment, plant.mass..g) 

################ ################ ################
# combine T1  T2  T3 T4 timepoints
################ ################ ################
O2.tanks<-rbind(T1.O2,T2.O2, T3.O2, T4.O2)
```
  
TIME POINT 1: Change in O2% and NEP and NER --   
make figures of the data for timepoint 1  
```{r Time1 O2}
########## 
Fig.formatting<-(theme_classic()) +
  theme(text=element_text(size=10),
        axis.line=element_blank(),
        legend.text.align = 0,
        legend.text=element_text(size=10),
        #legend.title = element_blank(),
        panel.border = element_rect(fill=NA, colour = "black", size=1),
        aspect.ratio=1, 
        axis.ticks.length=unit(0.25, "cm"),
        axis.text.y=element_text(
          margin=unit(c(0.5, 0.5, 0.5, 0.5), "cm"), colour="black", size=10), 
        axis.text.x=element_text(
          margin=unit(c(0.5, 0.5, 0.5, 0.5), "cm"), colour="black", size=8)) +
  theme(legend.key.size = unit(0.4, "cm")) +
  theme(aspect.ratio=1.3) +
  theme(panel.spacing=unit(c(0, 0, 0, 0), "cm"))

#########################################################
############################################################################
###############################################################################################
###################

# total oxygen % plot for the 3 time points (dawn-dusk-dawn)

T1.AM1<-ggplot(T1.O2, aes(x=plant.mass..g, y=dawn1, color=Treatment)) + 
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  geom_hline(yintercept=0, linetype="longdash", color = "gray") +
  ggtitle("T1.dawn1")+
  coord_cartesian(ylim=c(-30, 150)) +
  ylab(expression(paste("O"[2],"%"))) +
  xlab("plant material (g)") +
  Fig.formatting

T1.PM1<-ggplot(T1.O2, aes(x=plant.mass..g, y=dusk1, color=Treatment)) + 
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  geom_hline(yintercept=0, linetype="longdash", color = "gray") +
  ggtitle("T1.dusk1")+
  coord_cartesian(ylim=c(-30, 150)) +
  ylab(expression(paste("O"[2],"%"))) +
  xlab("plant material (g)") +
  Fig.formatting

T1.AM2<-ggplot(T1.O2, aes(x=plant.mass..g, y=dawn2, color=Treatment)) + 
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  geom_hline(yintercept=0, linetype="longdash", color = "gray") +
  ggtitle("T1.dawn2")+
  coord_cartesian(ylim=c(-30, 150)) +
  ylab(expression(paste("O"[2],"%"))) +
  xlab("plant material (g)") +
  Fig.formatting

extract.legend <- get_legend(
  # create some space to the left of the legend
  T1.AM1 + theme(legend.box.margin = margin(0, 0, 0, 10)))

O2.T1<-plot_grid(
  T1.AM1+ theme(legend.position = "none"), 
  T1.PM1+ theme(legend.position = "none"), 
  T1.AM2+ theme(legend.position = "none"),
  extract.legend, 
  rel_widths = c(8,8,8,3), ncol=4)

#ggsave("figures/O2.T1.pdf", height=5, width=9)

#################################
############################################
# net ecosystem production and respiration

#Time 1 NEP
NEP.T1<-ggplot(T1.O2, aes(x=plant.mass..g, y=NEP, color=Treatment)) + 
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  geom_hline(yintercept=0, linetype="longdash", color = "gray") +
  ylab(expression(paste("Net Production (", Delta, "O"[2],"%)"))) +
  coord_cartesian(ylim=c(-20, 50)) +
  xlab("plant material (g)")+
  Fig.formatting

#Time 1 NER
NER.T1<-ggplot(T1.O2, aes(x=plant.mass..g, y=NER, color=Treatment)) + 
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  geom_hline(yintercept=0, linetype="longdash", color = "gray") +
  coord_cartesian(ylim=c(-40, 10)) +
  ylab(expression(paste("Net Respiration (", Delta, "O"[2],"%)"))) +
  xlab("plant material (g)") +
  Fig.formatting


O2.change.T1<-plot_grid(
  NEP.T1+ theme(legend.position = "none") + ggtitle("Time 1"), 
  NER.T1+ theme(legend.position = "none") + ggtitle(""),
  extract.legend, 
  rel_widths = c(8,8,3), ncol=3)

#ggsave("figures/O2.change.T1.pdf", height=5, width=8)
```
  
TIME POINT 2: Change in O2% and NEP and NER --
make figures of the data for timepoint 2
```{r Time2 O2}
############################################################
##############################################################################
#############################################################################################
###################

# total oxygen % plot for the 3 time points (dawn-dusk-dawn)

T2.AM1<-ggplot(T2.O2, aes(x=plant.mass..g, y=dawn1, color=Treatment)) + 
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  geom_hline(yintercept=0, linetype="longdash", color = "gray") +
  ggtitle("T2.dawn1")+
  coord_cartesian(ylim=c(-30, 150)) +
  ylab(expression(paste("O"[2],"%"))) +
  xlab("plant material (g)") +
  Fig.formatting

T2.PM1<-ggplot(T2.O2, aes(x=plant.mass..g, y=dusk1, color=Treatment)) + 
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  geom_hline(yintercept=0, linetype="longdash", color = "gray") +
  ggtitle("T2.dusk1")+
  coord_cartesian(ylim=c(-30, 150)) +
  ylab(expression(paste("O"[2],"%"))) +
  xlab("plant material (g)") +
  Fig.formatting

T2.AM2<-ggplot(T2.O2, aes(x=plant.mass..g, y=dawn2, color=Treatment)) + 
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  geom_hline(yintercept=0, linetype="longdash", color = "gray") +
  ggtitle("T2.dawn2")+
  coord_cartesian(ylim=c(-30, 150)) +
  ylab(expression(paste("O"[2],"%"))) +
  xlab("plant material (g)") +
  Fig.formatting

O2.T2<-plot_grid(
  T2.AM1+ theme(legend.position = "none"), 
  T2.PM1+ theme(legend.position = "none"),
  T2.AM2+ theme(legend.position = "none"),
  extract.legend, 
  rel_widths = c(8,8,8,3), ncol=4)

#ggsave("figures/O2.T2.pdf", height=5, width=9)


############################################################
##############################################################################################
# net ecosystem production and respiration

#Time 2 NEP
NEP.T2<-ggplot(T2.O2, aes(x=plant.mass..g, y=NEP, color=Treatment)) + 
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  geom_hline(yintercept=0, linetype="longdash", color = "gray") +
  coord_cartesian(ylim=c(-20, 50)) +
  ylab(expression(paste("Net Production (", Delta, "O"[2],"%)"))) +
  xlab("plant material (g)") +
  Fig.formatting

#Time 2 NER
NER.T2<-ggplot(T2.O2, aes(x=plant.mass..g, y=NER, color=Treatment)) + 
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment),alpha=0.1) +
  geom_hline(yintercept=0, linetype="longdash", color = "gray") +
  coord_cartesian(ylim=c(-40, 10)) +
  ylab(expression(paste("Net Respiration (", Delta, "O"[2],"%)"))) +
  xlab("plant material (g)") +
  Fig.formatting

O2.change.T2<-plot_grid(
  NEP.T2+ theme(legend.position = "none") + ggtitle("Time 2"),
  NER.T2+ theme(legend.position = "none") + ggtitle(""),
  extract.legend, 
  rel_widths = c(8,8,3), ncol=3)

#ggsave("figures/O2.change.T2.pdf", height=5, width=8)


```

TIME POINT 3: Change in O2% and NEP and NER --
make figures of the data for timepoint 3
```{r Time3 O2}
############################################################
##############################################################################
#############################################################################################
###################

# total oxygen % plot for the 3 time points (dawn-dusk-dawn)

T3.AM1<-ggplot(T3.O2, aes(x=plant.mass..g, y=dawn1, color=Treatment)) + 
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  geom_hline(yintercept=0, linetype="longdash", color = "gray") +
  ggtitle("T3.dawn1")+
  coord_cartesian(ylim=c(-30, 150)) +
  ylab(expression(paste("O"[2],"%"))) +
  xlab("plant material (g)") +
  Fig.formatting

T3.PM1<-ggplot(T3.O2, aes(x=plant.mass..g, y=dusk1, color=Treatment)) + 
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  geom_hline(yintercept=0, linetype="longdash", color = "gray") +
  ggtitle("T3.dusk1")+
  coord_cartesian(ylim=c(-30, 150)) +
  ylab(expression(paste("O"[2],"%"))) +
  xlab("plant material (g)") +
  Fig.formatting

T3.AM2<-ggplot(T3.O2, aes(x=plant.mass..g, y=dawn2, color=Treatment)) + 
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  geom_hline(yintercept=0, linetype="longdash", color = "gray") +
  ggtitle("T3.dawn2")+
  coord_cartesian(ylim=c(-30, 150)) +
  ylab(expression(paste("O"[2],"%"))) +
  xlab("plant material (g)") +
  Fig.formatting

O2.T3<-plot_grid(
  T3.AM1+ theme(legend.position = "none"), 
  T3.PM1+ theme(legend.position = "none"),
  T3.AM2+ theme(legend.position = "none"),
  extract.legend, 
  rel_widths = c(8,8,8,3), ncol=4)

#ggsave("figures/O2.T3.pdf", height=5, width=9)


############################################################
##############################################################################################
# net ecosystem production and respiration

#Time 3 NEP
NEP.T3<-ggplot(T3.O2, aes(x=plant.mass..g, y=NEP, color=Treatment)) + 
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  geom_hline(yintercept=0, linetype="longdash", color = "gray") +
  coord_cartesian(ylim=c(-20, 50)) +
  ylab(expression(paste("Net Production (", Delta, "O"[2],"%)"))) +
  xlab("plant material (g)") +
  Fig.formatting

#Time 3 NER
NER.T3<-ggplot(T3.O2, aes(x=plant.mass..g, y=NER, color=Treatment)) + 
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment),alpha=0.1) +
  geom_hline(yintercept=0, linetype="longdash", color = "gray") +
  coord_cartesian(ylim=c(-40, 10)) +
  ylab(expression(paste("Net Respiration (", Delta, "O"[2],"%)"))) +
  xlab("plant material (g)") +
  Fig.formatting

O2.change.T3<-plot_grid(
  NEP.T3+ theme(legend.position = "none") + ggtitle("Time 3"),
  NER.T3+ theme(legend.position = "none") + ggtitle(""),
  extract.legend, 
  rel_widths = c(8,8,3), ncol=3)

#ggsave("figures/O2.change.T3.pdf", height=5, width=8)


```
  
TIME POINT 4: Change in O2% and NEP and NER --
make figures of the data for timepoint 4
```{r Time4 O2}
############################################################
##############################################################################
#############################################################################################
###################

# total oxygen % plot for the 3 time points (dawn-dusk-dawn)

T4.AM1<-ggplot(T4.O2, aes(x=plant.mass..g, y=dawn1, color=Treatment)) + 
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  geom_hline(yintercept=0, linetype="longdash", color = "gray") +
  ggtitle("T4.dawn1")+
  coord_cartesian(ylim=c(-30, 150)) +
  ylab(expression(paste("O"[2],"%"))) +
  xlab("plant material (g)") +
  Fig.formatting

T4.PM1<-ggplot(T4.O2, aes(x=plant.mass..g, y=dusk1, color=Treatment)) + 
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  geom_hline(yintercept=0, linetype="longdash", color = "gray") +
  ggtitle("T4.dusk1")+
  coord_cartesian(ylim=c(-30, 150)) +
  ylab(expression(paste("O"[2],"%"))) +
  xlab("plant material (g)") +
  Fig.formatting

T4.AM2<-ggplot(T4.O2, aes(x=plant.mass..g, y=dawn2, color=Treatment)) + 
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  geom_hline(yintercept=0, linetype="longdash", color = "gray") +
  ggtitle("T4.dawn2")+
  coord_cartesian(ylim=c(-30, 150)) +
  ylab(expression(paste("O"[2],"%"))) +
  xlab("plant material (g)") +
  Fig.formatting

O2.T4<-plot_grid(
  T4.AM1+ theme(legend.position = "none"), 
  T4.PM1+ theme(legend.position = "none"),
  T4.AM2+ theme(legend.position = "none"),
  extract.legend, 
  rel_widths = c(8,8,8,3), ncol=4)

#ggsave("figures/O2.T4.pdf", height=5, width=9)


############################################################
##############################################################################################
# net ecosystem production and respiration

#Time 4 NEP
NEP.T4<-ggplot(T4.O2, aes(x=plant.mass..g, y=NEP, color=Treatment)) + 
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  geom_hline(yintercept=0, linetype="longdash", color = "gray") +
  coord_cartesian(ylim=c(-20, 50)) +
  ylab(expression(paste("Net Production (", Delta, "O"[2],"%)"))) +
  xlab("plant material (g)") +
  Fig.formatting

#Time 4 NER
NER.T4<-ggplot(T4.O2, aes(x=plant.mass..g, y=NER, color=Treatment)) + 
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment),alpha=0.1) +
  geom_hline(yintercept=0, linetype="longdash", color = "gray") +
  coord_cartesian(ylim=c(-40, 10)) +
  ylab(expression(paste("Net Respiration (", Delta, "O"[2],"%)"))) +
  xlab("plant material (g)") +
  Fig.formatting

O2.change.T4<-plot_grid(
  NEP.T4+ theme(legend.position = "none") + ggtitle("Time 4"),
  NER.T4+ theme(legend.position = "none") + ggtitle(""),
  extract.legend, 
  rel_widths = c(8,8,3), ncol=3)

#ggsave("figures/O2.change.T4.pdf", height=5, width=8)
```


```{r NEP NER compile}
########## ########## ########## ########## ########## 
########## compile all figs

##### All dawn-dusk-dawn measurements
O2change.alltimes<-plot_grid(
  O2.T1,
  O2.T2,
  O2.T3,
  O2.T4, ncol=1)

ggsave("figures/O2change.alltimes.pdf", height=12, width=10)


##### All photosynthesis and respiration

PnR.alltimes.long<-plot_grid(
  NEP.T1+ theme(legend.position = "none")+ ggtitle("Time-1"),
  NEP.T2+ theme(legend.position = "none")+ ggtitle("Time-2"),
  NEP.T3+ theme(legend.position = "none")+ ggtitle("Time-3"),
  NEP.T4+ theme(legend.position = "none")+ ggtitle("Time-4"), extract.legend,
  NER.T1+ theme(legend.position = "none")+ ggtitle(""),
  NER.T2+ theme(legend.position = "none")+ ggtitle(""),
  NER.T3+ theme(legend.position = "none")+ ggtitle(""),
  NER.T4+ theme(legend.position = "none")+ ggtitle(""), extract.legend,
  rel_widths = c(8,8,8,8,3, 8,8,8,8,3), ncol=5)

ggsave("figures/PnR.alltimes.long.pdf", height=6, width=14)


# alternative plot
PnR.alltimes<-plot_grid(
  O2.change.T1,
  O2.change.T2,
  O2.change.T3,
  O2.change.T4, ncol=1)

ggsave("figures/PnR.alltimes.pdf", height=12, width=8)


```
  
  
###Chlorophyll  
import 
```{r chlorophyll}
## import treatment IDs
chl.dat<- read.csv("data/Pyro_chlorophyll.csv")  
chl.dat<-chl.dat[!(chl.dat$Time.point=="First.round"),] # remove first round from aborted mission

#remove outlier
chl.out.rem<-chl.dat[!(chl.dat$chla..ug.L>=10),]


#plot

chla.T0<-ggplot(chl.dat[(chl.dat$Time.point=="T0"),], aes(x=plant.mass..g, y=chla..ug.L, color=Treatment)) +
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  coord_cartesian(ylim=c(0, 20)) +
  ggtitle("Time-0") +
  ylab(expression(paste("chlorophyll", ~(mu*g~L^-1), sep=""))) +
  xlab("plant material (g)") +
  Fig.formatting


chla.T1<-ggplot(chl.dat[(chl.dat$Time.point=="T1"),], aes(x=plant.mass..g, y=chla..ug.L, color=Treatment)) +
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  coord_cartesian(ylim=c(0, 20)) +
  ggtitle("Time-1") +
  ylab(expression(paste("chlorophyll", ~(mu*g~L^-1), sep=""))) +
  xlab("plant material (g)") +
  Fig.formatting

chla.T2<-ggplot(chl.dat[(chl.dat$Time.point=="T2"),], aes(x=plant.mass..g, y=chla..ug.L, color=Treatment)) +
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  coord_cartesian(ylim=c(0, 20)) +
  ggtitle("Time-2") +
  ylab(expression(paste("chlorophyll", ~(mu*g~L^-1), sep=""))) +
  xlab("plant material (g)") +
  Fig.formatting

chla.T3<-ggplot(chl.dat[(chl.dat$Time.point=="T3"),], aes(x=plant.mass..g, y=chla..ug.L, color=Treatment)) +
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  coord_cartesian(ylim=c(0, 20)) +
  ggtitle("Time-3") +
  ylab(expression(paste("chlorophyll", ~(mu*g~L^-1), sep=""))) +
  xlab("plant material (g)") +
  Fig.formatting

chla.T4<-ggplot(chl.dat[(chl.dat$Time.point=="T4"),], aes(x=plant.mass..g, y=chla..ug.L, color=Treatment)) +
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  coord_cartesian(ylim=c(0, 20)) +
  ggtitle("Time-4") +
  ylab(expression(paste("chlorophyll", ~(mu*g~L^-1), sep=""))) +
  xlab("plant material (g)") +
  Fig.formatting


chl.plots<-plot_grid(
  chla.T0+ theme(legend.position = "none"),
  chla.T1+ theme(legend.position = "none"),
  chla.T2+ theme(legend.position = "none"),
  chla.T3+ theme(legend.position = "none"),
  chla.T4+ theme(legend.position = "none"),
  extract.legend, 
  rel_widths = c(8,8,8,8,8,3), ncol=6)


chl.plots

ggsave("figures/chla.alltime.pdf", height=5, width=15)



```
  
  
###DOC  
import and do a loop to clean up all files and make stacked data in single df
```{r DOC}
## import treatment IDs
IDs<-read.csv("data/treatment.IDs.csv")

##### grab files in a list
Total.DOC.files <- list.files(path="data/DOC.TN", pattern = "csv$", full.names = T)
Total.DOC.files

##### what are the file names, sans extensions using package 'tools'
file.names<-file_path_sans_ext(list.files(path="data/DOC.TN", pattern = "csv$", full.names = F))
file.names
############ formatting all data in for loop
  for(i in 1:length(Total.DOC.files))
    {
  data<-read.csv(Total.DOC.files[i], sep=",")
  data<-data[,c(1,3,4)] # removed columns we don't need
  data$File<-Total.DOC.files[i]
  colnames(data)<-c("Tank", "DOC..mg.L", "TN..mg.L", "File")
  data$Tank<- IDs$Tank
  data$Tank<-as.numeric(as.character(data$Tank)) # make the column of samples all numeric
  data <- data[!is.na(as.numeric(as.character(data$Tank))),] # remove all rows that aren't numeric/tanks
  data$Treatment<-IDs$Treatment
  data$plant.mass..g<-IDs$plant.mass..g
  make.names(assign(paste(file.names[i], sep=""), data)) # make the file name the name of new df for loop df
  }
########## this is the end of the loop


ls() #see all dfs you've made, the above will be df matching their file names

#Combine files from loop to single df
DOC.df<-rbind(DOC_T0, DOC_T1, DOC_T2, DOC_T3, DOC_T4)

DOC.df$File <- sapply(strsplit(DOC.df$File, "/"), `[`, 3) # extract sample names

# alternative way to code the above
#give the 10th-24th character of the file name, removing the rest
#DOC.df$File<-substr(DOC.df$File, 13, 27) 

#alternatively
# remove the 9 letters ('^.) at start 
# remove the 4 letters (.$') at end
#DOC.df$File<-gsub('^.........|....$', '', DOC.df$File) 

# if equals DOC_T0_11052021 then, T0, if not then T1
DOC.df$Time.point<- as.factor(ifelse(DOC.df$File=="DOC_T0.csv", "T0",
         ifelse (DOC.df$File=="DOC_T1.csv", "T1",
           ifelse (DOC.df$File=="DOC_T2.csv", "T2", 
                   ifelse(DOC.df$File=="DOC_T3.csv", "T3", "T4")))))

#rearrange
DOC.df<- DOC.df %>% 
  select(File, Time.point, Treatment, Tank, plant.mass..g, DOC..mg.L, TN..mg.L) 

```

DOC plots
```{r DOC plots}
DOC.T0<-ggplot(DOC.df[(DOC.df$Time.point=="T0"),], aes(x=plant.mass..g, y=DOC..mg.L, color=Treatment)) +
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  coord_cartesian(ylim=c(0, 60)) +
  ggtitle("Time-0") +
  ylab("DOC (mg/L)") +
  xlab("plant material (g)") +
  Fig.formatting

DOC.T1<-ggplot(DOC.df[(DOC.df$Time.point=="T1"),], aes(x=plant.mass..g, y=DOC..mg.L, color=Treatment)) + 
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  coord_cartesian(ylim=c(0, 60)) +
  ggtitle("Time-1") +
  ylab("DOC (mg/L)") +
  xlab("plant material (g)") +
  Fig.formatting
  
DOC.T2<-ggplot(DOC.df[(DOC.df$Time.point=="T2"),], aes(x=plant.mass..g, y=DOC..mg.L, color=Treatment)) + 
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  coord_cartesian(ylim=c(0, 60)) +
  ggtitle("Time-2") +
  ylab("DOC (mg/L)") +
  xlab("plant material (g)") +
  Fig.formatting

  
DOC.T3<-ggplot(DOC.df[(DOC.df$Time.point=="T3"),], aes(x=plant.mass..g, y=DOC..mg.L, color=Treatment)) + 
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  coord_cartesian(ylim=c(0, 60)) +
  ggtitle("Time-3") +
  ylab("DOC (mg/L)") +
  xlab("plant material (g)") +
  Fig.formatting

DOC.T4<-ggplot(DOC.df[(DOC.df$Time.point=="T4"),], aes(x=plant.mass..g, y=DOC..mg.L, color=Treatment)) + 
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  coord_cartesian(ylim=c(0, 60)) +
  ggtitle("Time-4") +
  ylab("DOC (mg/L)") +
  xlab("plant material (g)") +
  Fig.formatting

DOC.plots<-plot_grid(
  DOC.T0+ theme(legend.position = "none"),
  DOC.T1+ theme(legend.position = "none"),
  DOC.T2+ theme(legend.position = "none"),
  DOC.T3+ theme(legend.position = "none"),
  DOC.T4+ theme(legend.position = "none"),
  extract.legend, 
  rel_widths = c(8,8,8,8,8, 3), ncol=6)


DOC.plots

ggsave("figures/DOC.alltime.pdf", height=5, width=15)

```

Total N plots
```{r TN plots}
TN.T0<-ggplot(DOC.df[(DOC.df$Time.point=="T0"),], aes(x=plant.mass..g, y=TN..mg.L, color=Treatment)) + 
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  coord_cartesian(ylim=c(0, 2)) +
  ggtitle("Time-0") +
  ylab("TN (mg/L)") +
  xlab("plant material (g)") +
  Fig.formatting
 
TN.T1<-ggplot(DOC.df[(DOC.df$Time.point=="T1"),], aes(x=plant.mass..g, y=TN..mg.L, color=Treatment)) + 
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  coord_cartesian(ylim=c(0, 2)) +
  ggtitle("Time-1") +
  ylab("TN (mg/L)") +
  xlab("plant material (g)") +
  Fig.formatting

TN.T2<-ggplot(DOC.df[(DOC.df$Time.point=="T2"),], aes(x=plant.mass..g, y=TN..mg.L, color=Treatment)) + 
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  coord_cartesian(ylim=c(0, 2)) +
  ggtitle("Time-2") +
  ylab("TN (mg/L)") +
  xlab("plant material (g)") +
  Fig.formatting

TN.T3<-ggplot(DOC.df[(DOC.df$Time.point=="T3"),], aes(x=plant.mass..g, y=TN..mg.L, color=Treatment)) + 
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  coord_cartesian(ylim=c(0, 2)) +
  ggtitle("Time-3") +
  ylab("TN (mg/L)") +
  xlab("plant material (g)") +
  Fig.formatting

TN.T4<-ggplot(DOC.df[(DOC.df$Time.point=="T4"),], aes(x=plant.mass..g, y=TN..mg.L, color=Treatment)) + 
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  coord_cartesian(ylim=c(0, 2)) +
  ggtitle("Time-4") +
  ylab("TN (mg/L)") +
  xlab("plant material (g)") +
  Fig.formatting
  
TN.plots<-plot_grid(
  TN.T0+ theme(legend.position = "none"),
  TN.T1+ theme(legend.position = "none"),
  TN.T2+ theme(legend.position = "none"),
  TN.T3+ theme(legend.position = "none"),
  TN.T4+ theme(legend.position = "none"),
  extract.legend, 
  rel_widths = c(8,8,8,8,8,3), ncol=6)

ggsave("figures/TN.alltime.pdf", height=5, width=15)

```


### Isotopes
Time 1 isotopes
```{r isotopes d15N}

#########  Time 1
topes<-read.csv("data/Isotopes/Pyro_isotopes.csv")
topes$C.N <-(topes$Total.C..ug/12)/(topes$Total.N..ug/14) # C mol : N mol

cols<-c("Time.point", "Treatment", "Type", "Tank") # columns to make factors
topes[cols] <- lapply(topes[cols], factor) # make all these factors

##### make data frames
# treatment data df
topes.trt<-topes[(topes$Treatment=="burned" | topes$Treatment=="unburned"),]

# control and start plant materials df
topes.controls<-topes[!(topes$Treatment=="burned" | topes$Treatment=="unburned"),]

#### isotope plots for treatments

iso.plot.T1<-ggplot(topes.trt[(topes.trt$Time.point=="T1"),], 
                 aes(x=plant.mass..g, y=d15N, color=Treatment, linetype=Type)) +
  geom_point(aes(shape=Type)) +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  ylab(expression(paste(delta^{15}, N, " (\u2030, air)"))) +
  xlab("plant material (g)") +
  coord_cartesian(ylim=c(0, 250)) +
  Fig.formatting +
  theme(legend.key.size = unit(2,"line"))

############ Time 2

#### isotope plots for treatments
iso.plot.T2<-ggplot(topes.trt[(topes.trt$Time.point=="T2"),],
                 aes(x=plant.mass..g, y=d15N, color=Treatment, linetype=Type)) +
  geom_point(aes(shape=Type)) +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  ylab(expression(paste(delta^{15}, N, " (\u2030, air)"))) +
  xlab("plant material (g)") +
  coord_cartesian(ylim=c(0, 250)) +
  Fig.formatting +
  theme(legend.key.size = unit(2,"line"))

##########
extract.legend.iso <- get_legend(
  # create some space to the left of the legend
  iso.plot.T1 + theme(legend.box.margin = margin(0, 0, 0, 10)))

###### group plot
isotope.plots<-plot_grid(
  iso.plot.T1+ theme(legend.position = "none") + ggtitle("Time-1"),
  iso.plot.T2+ theme(legend.position = "none") + ggtitle("Time-2") ,
  extract.legend.iso, 
  rel_widths = c(8,8,3), ncol=3)

ggsave("figures/Isotope.d15N.pdf", height=5, width=8, encod="MacRoman")
```


C:N ratio for plankton
```{r isotopes CN}
############# C vs N plot

### Time 1 C vs. N
iso.plot.CN.T1<-ggplot(topes.trt[(topes.trt$Time.point=="T1"),],
                    aes(x=Total.C..ug, y=Total.N..ug, color=Treatment)) +
  geom_point(aes(shape=Type)) +
  ggtitle("Time-1")+
  scale_shape_manual(name="Plankton", values = c(17, 16), 
                     labels = c(expression(paste("> 63"~mu,"m")), 
                                expression(paste("< 63"~mu,"m")))) +
  scale_color_manual(name="Treatments", values = c("brown1", "mediumseagreen")) +
  # having this repeated below for some reason hacks the legend in a positive way, so leave it in.
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  ylim(c(0,550)) +
  xlim(c(0,3000)) +
  ylab("Total N (ug)") +
  xlab("Total C (ug)") +
  Fig.formatting +
  theme(legend.key.size = unit(2,"line"))


### Time 2 C vs. N
iso.plot.CN.T2<-ggplot(topes.trt[(topes.trt$Time.point=="T2"),],
                    aes(x=Total.C..ug, y=Total.N..ug, color=Treatment)) +
  geom_point(aes(shape=Type)) +
  ggtitle("Time-2")+
  scale_shape_manual(name="Plankton", values = c(17, 16), 
                     labels = c(expression(paste("> 63"~mu,"m")), 
                                expression(paste("< 63"~mu,"m")))) +
  scale_color_manual(name="Treatments", values = c("brown1", "mediumseagreen")) +
  scale_color_manual(values = c("brown1", "mediumseagreen")) +
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1, show.legend = FALSE) +
  ylim(c(0,550)) +
  xlim(c(0,3000)) +
  ylab("Total N (ug)") +
  xlab("Total C (ug)") +
  Fig.formatting +
  theme(legend.key.size = unit(2,"line"))


######### C:N ratio plot by biomass
iso.plot.CNrat.T1<-ggplot(topes.trt[(topes.trt$Time.point=="T1"),], 
                         aes(y=C.N, x=plant.mass..g, color=Treatment)) +
  geom_point(aes(shape=Type)) +
  ggtitle("Time-1")+ 
  coord_cartesian(ylim=c(0, 20)) +
  scale_shape_manual(name="Plankton", values = c(17, 16), 
                     labels = c(expression(paste("> 63"~mu,"m")), 
                                expression(paste("< 63"~mu,"m")))) +
  scale_color_manual(name="Treatments", values = c("brown1", "mediumseagreen")) +
  scale_color_manual(values = c("brown1", "mediumseagreen")) +
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1, show.legend = FALSE) +
  ylab("C:N") +
  xlab("plant material (g)") +
  Fig.formatting +
  theme(legend.key.size = unit(2,"line"))


######## Time 2
iso.plot.CNrat.T2<-ggplot(topes.trt[(topes.trt$Time.point=="T2"),], 
                         aes(y=C.N, x=plant.mass..g, color=Treatment)) +
  geom_point(aes(shape=Type)) +
  ggtitle("Time-2")+ 
  coord_cartesian(ylim=c(0, 20)) +
  scale_shape_manual(name="Plankton", values = c(17, 16), 
                     labels = c(expression(paste("> 63"~mu,"m")), 
                                expression(paste("< 63"~mu,"m")))) +
  scale_color_manual(name="Treatments", values = c("brown1", "mediumseagreen")) +
  scale_color_manual(values = c("brown1", "mediumseagreen")) +
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1, show.legend = FALSE) +
  ylab("C:N") +
  xlab("plant material (g)") +
  Fig.formatting +
  theme(legend.key.size = unit(2,"line"))


### get legend
extract.legend.iso2 <- get_legend(
  # create some space to the left of the legend
  iso.plot.CN.T1 + theme(legend.box.margin = margin(0, 0, 0, 10)))

## combine plots
isoCNplot<- plot_grid(iso.plot.CN.T1 + theme(legend.position = "none"), 
                    iso.plot.CN.T2 + theme(legend.position = "none"),
                    iso.plot.CNrat.T1 + theme(legend.position = "none"),
                    iso.plot.CNrat.T2 + theme(legend.position = "none"),
                    extract.legend.iso2, rel_widths = c(8,8,8,8,3), ncol=5)

isoCNplot
ggsave("figures/iso.plot.CN.alltime.pdf", encod="MacRoman", height=6, width=15)
```


```{r control isotopes}
########## ########## ########## 
########## control plots

# set levels
topes.controls$Type<-factor(topes.controls$Type, 
                         levels=c("blank", "plankton.stock", 
                                  "willow.unburned", "willow.burned",
                                  "sage.unburned", "sage.burned", 
                                  "sage.veryburned", "sage.stem.burned"))


#### controls d15N
iso.plot.control.d15N<-ggplot(data=topes.controls,  aes(x=Type, y=d15N, fill=Type)) +
  geom_boxplot() +
  geom_point(pch = 21, position = position_jitterdodge(), alpha=0.6)+
  ylab(expression(paste(delta^{15}, N, " (\u2030, air)"))) +
  scale_fill_manual(values = c("azure2", "cornflowerblue",
                                "darkgoldenrod1", "darkgoldenrod",
                                "darkseagreen3", "darkolivegreen3",
                                "darkolivegreen4", "darkgreen")) +
  xlab("control types") +  Fig.formatting +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 


###### control C:N
topes.controls.CN<-topes.controls %>% drop_na(C.N) # drop the NAs for C.N, makes plotting problematic

iso.plot.control.CN<-ggplot(data=topes.controls.CN,  aes(x=Type, y=C.N, fill=Type)) +
  geom_boxplot() +
  geom_point(pch = 21, position = position_jitterdodge(), alpha=0.6)+
  ylab("C:N") +
  scale_fill_manual(values = c("cornflowerblue",
                                "darkgoldenrod1", "darkgoldenrod",
                                "darkseagreen3", "darkolivegreen3",
                                "darkolivegreen4", "darkgreen")) +
  xlab("control types") +  Fig.formatting +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))


####### combine plots
# legend
extract.legend.cont <- get_legend(
  # create some space to the left of the legend
  iso.plot.control.d15N + theme(legend.box.margin = margin(0, 0, 0, 10)))


## combine 
control.iso.alltime<- plot_grid(iso.plot.control.d15N + theme(legend.position = "none"), 
                    iso.plot.control.CN + theme(legend.position = "none"),
                    extract.legend.cont, rel_widths = c(8,8,3), ncol=3)

ggsave("figures/iso.controls.pdf", encod="MacRoman", height=5, width=10)
```

Use a mixing model to calcualte % sage
```{r percent sage}
# mixing model
head(topes.trt)
topes.trt<-droplevels(topes.trt)

d15N.cont<-aggregate(d15N~Type, topes.controls, FUN=mean)
d15N.cont.n<-aggregate(d15N~Type, topes.controls, FUN=length)

d15N.cont # summary mean d15N
sage.mean<-d15N.cont[c(5:8),]
willow.mean<-d15N.cont[c(3:4),]
mean(sage.mean$d15N) # sage ~298
mean(willow.mean$d15N) #  willow 13

# summary mean atom percent enrichment
F.cont<-aggregate(at.P..15N  ~ Type, topes.controls, mean)
sage.meanF<-F.cont[c(5:8),]
willow.meanF<-F.cont[c(3:4),]
mean(sage.meanF$at.P..15N) # sage ~0.475
mean(willow.meanF$at.P..15N) # willow 0.371

# 2 source mixing model (Post 2002)

# alpha = percent Sage from food web 1
# %Sage = (d15N sample - d15N base 2 [i.e., no-label food])/ (d15N sage food 1 - source 2)
#  d15N values of base 2 = 11 permil for algae/plankton stock 
#  d15N value of base 1 = 298 permil for sage

topes.trt$percent.sage<-(topes.trt$d15N-11)/(298-11)

# unicode text for micrometer = \u03BC, use this in legend

iso.mix.T1<-ggplot(topes.trt[(topes.trt$Time.point=="T1"),], aes(x=plant.mass..g, y=100*percent.sage, color=Treatment)) +
  geom_point(aes(shape=Type)) +
  scale_shape_manual(name="Plankton", values = c(17, 16), 
                     labels = c(expression(paste("> 63"~mu,"m")), 
                                expression(paste("< 63"~mu,"m")))) +
  scale_color_manual(name="Treatments", values = c("brown1", "mediumseagreen")) +
  geom_smooth(method = "gam", formula = y ~ s(x, bs = "tp", k = 5),
              aes(fill=Treatment), alpha=0.1, show.legend = FALSE) +
  ylab("% Sage")+
  xlab("plant material (g)") +
  ggtitle("Time-1") +
  coord_cartesian(ylim=c(0, 100)) +
  Fig.formatting +
  theme(legend.key.size = unit(1,"line"))


######################

iso.mix.T2<-ggplot(topes.trt[(topes.trt$Time.point=="T2"),], aes(x=plant.mass..g, y=100*percent.sage, color=Treatment)) +
  geom_point(aes(shape=Type)) +
  scale_shape_manual(name="Plankton", values = c(17, 16), 
                     labels = c(expression(paste("> 63"~mu,"m")), 
                                expression(paste("< 63"~mu,"m")))) +
  scale_color_manual(name="Treatments", values = c("brown1", "mediumseagreen")) +
  geom_smooth(method = "gam", formula = y ~ s(x, bs = "tp", k = 5),
              aes(fill=Treatment), alpha=0.1, show.legend = FALSE) +
  ylab("% Sage")+
  xlab("plant material (g)") +
  ggtitle("Time-2") +
  coord_cartesian(ylim=c(0, 100)) +
  Fig.formatting +
  theme(legend.key.size = unit(1,"line"))


# legend
extract.legend.mix <- get_legend(
  # create some space to the left of the legend
  iso.mix.T2 + theme(legend.box.margin = margin(0, 0, 0, 10)))


## combine 
control.iso.alltime<- plot_grid(iso.mix.T1 + theme(legend.position = "none"), 
                    iso.mix.T2 + theme(legend.position = "none"),
                    extract.legend.mix, rel_widths = c(8,8,3), ncol=3)

ggsave("figures/Isotope.mixmodel.pdf", encod="MacRoman", height=5, width=10)



### some models



############# all plankton T1
mod1c <- gam(percent.sage ~ s(plant.mass..g, k=5, bs="tp")+
            s(plant.mass..g, Treatment, k=5, bs="re")+
            s(plant.mass..g, Type, k=2, bs="re"),
            data=topes.trt, method="REML", family="gaussian")

mod2c <- gam(percent.sage ~ s(plant.mass..g, k=5, bs="tp")+
            s(plant.mass..g, Treatment, k=5, bs="re"),
            data=topes.trt, method="REML", family="gaussian")

# model with biomass
mod3c <- gam(percent.sage ~ s(plant.mass..g, k=5, bs="tp"), 
            data=topes.trt, method="REML", family="gaussian")

AIC(mod1c, mod2c, mod3c) 
# mod1c preferred
# compare across

summary(mod1c)
anova(mod1c)



############# all plankton T2
mod1c <- gam(percent.sage ~ s(plant.mass..g, k=5, bs="tp")+
            s(plant.mass..g, Treatment, k=5, bs="re")+
            s(plant.mass..g, Type, k=2, bs="re"),
            data=topes2.trt, method="REML", family="gaussian")

mod2c <- gam(percent.sage ~ s(plant.mass..g, k=5, bs="tp")+
            s(plant.mass..g, Treatment, k=5, bs="re"),
            data=topes2.trt, method="REML", family="gaussian")

# model with biomass
mod3c <- gam(percent.sage ~ s(plant.mass..g, k=5, bs="tp"), 
            data=topes2.trt, method="REML", family="gaussian")

AIC(mod1c, mod2c, mod3c) 
# mod1c preferred
# compare across

summary(mod1c)
anova(mod1c)






```



