---
title: "Pyromania"
author: "C Wall"
date: "1/4/2021"
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 4
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '4'
---

```{r setup chunk, setup, include = FALSE, cache=FALSE, message=FALSE, warning=FALSE}
if (!require('knitr')) install.packages('knitr'); library('knitr')
knitr::opts_chunk$set(warning=FALSE, message=FALSE, fig.align='center')

# load packages
if (!require("pacman")) install.packages("pacman") # for rapid install if not in library

# use pacman to load all the packages you are missing!
pacman::p_load('knitr', 'lme4', 'lmerTest', 'tidyverse', 'magrittr', 'effects', 'plyr', 'dplyr', 
               'plotrix', 'car',"gridExtra", "cowplot", "tools")


```


###YSI  
Time 1 and Time 2: import YSI data and produce plots of changes in O2% and net ecosystem productivity (NEP) and net ecosystem respiration (NER)  
```{r}
#load YSI data
YSI<-read.csv("data/Pyro_YSI.csv")

# fix date
YSI$Date<-as.character(YSI$Date)
YSI$Date<-as.POSIXct(YSI$Date, format="%m/%d/%Y")
YSI$Date<-as.Date(YSI$Date, format="%m/%d/%Y")


####### Time 1 change in O2 ################

#separate time points
YSI.T1<- YSI[(YSI$Time.point=="T1"),]

#calculate NEP for T1
T1.Prod<-YSI.T1[(YSI.T1$Date == "2021-11-15"),] # dawn and dusk for 12h period
T1.Dawn1<-T1.Prod[(T1.Prod$Dawn..Dusk == "dawn"),] # dawn-1 measurements
T1.Dusk<-T1.Prod[(T1.Prod$Dawn..Dusk == "dusk"),] # dusk measurements

T1.Dawn2<-YSI.T1[(YSI.T1$Date == "2021-11-16"),] # dawn-2 measurements, following AM

# make new dataframe
T1.O2<-(T1.Dawn1[,c(2,4:6)]) 
T1.O2$dawn1<-T1.Dawn1$DO.percent
T1.O2$dusk1<-T1.Dusk$DO.percent
T1.O2$dawn2<-T1.Dawn2$DO.percent

# NER = dusk - dawn (PM to AM, O2 change of day 1)
# NEP = dusk - dawn (PM to AM, O2 change of day 2)

T1.O2<- mutate(T1.O2, 
                NEP=dusk1 - dawn1,
                NER=dawn2 - dusk1) 

#sort
T1.O2<-T1.O2 %>% 
  arrange(Treatment, plant.mass..g) 


################ ################ ################
####### Time 2 change in O2 ################

#separate time points
YSI.T2<- YSI[(YSI$Time.point=="T2"),]

#calculate NEP for T2
T2.Prod<-YSI.T2[(YSI.T2$Date == "2021-12-06"),] # dawn and dusk for 12h period
T2.Dawn1<-T2.Prod[(T2.Prod$Dawn..Dusk == "dawn"),] # dawn-1 measurements
T2.Dusk<-T2.Prod[(T2.Prod$Dawn..Dusk == "dusk"),] # dusk measurements

T2.Dawn2<-YSI.T2[(YSI.T2$Date == "2021-12-07"),] # dawn-2 measurements, following AM

# make new dataframe
T2.O2<-(T2.Dawn1[,c(2,4:6)]) 
T2.O2$dawn1<-T2.Dawn1$DO.percent
T2.O2$dusk1<-T2.Dusk$DO.percent
T2.O2$dawn2<-T2.Dawn2$DO.percent

T2.O2<- mutate(T2.O2, 
                NEP=dusk1 - dawn1,
                NER=dawn2 - dusk1) 

#sort
T2.O2<-T2.O2 %>% 
  arrange(Treatment, plant.mass..g) 


################ ################ ################
####### Time 3 change in O2 ################

#separate time points
YSI.T3<- YSI[(YSI$Time.point=="T3"),]

#calculate NEP for T3
T3.Prod<-YSI.T3[(YSI.T3$Date == "2022-01-03"),] # dawn and dusk for 12h period
T3.Dawn1<-T3.Prod[(T3.Prod$Dawn..Dusk == "dawn"),] # dawn-1 measurements
T3.Dusk<-T3.Prod[(T3.Prod$Dawn..Dusk == "dusk"),] # dusk measurements

T3.Dawn2<-YSI.T3[(YSI.T3$Date == "2022-01-04"),] # dawn-2 measurements, following AM

# make new dataframe
T3.O2<-(T3.Dawn1[,c(2,4:6)]) 
T3.O2$dawn1<-T3.Dawn1$DO.percent
T3.O2$dusk1<-T3.Dusk$DO.percent
T3.O2$dawn2<-T3.Dawn2$DO.percent

T3.O2<- mutate(T3.O2, 
                NEP=dusk1 - dawn1,
                NER=dawn2 - dusk1) 

#sort
T3.O2<-T3.O2 %>% 
  arrange(Treatment, plant.mass..g) 

################ ################ ################
####### Time 3 change in O2 ################

#separate time points
YSI.T4<- YSI[(YSI$Time.point=="T4"),]

#calculate NEP for T4
T4.Prod<-YSI.T4[(YSI.T4$Date == "2022-02-02"),] # dawn and dusk for 12h period
T4.Dawn1<-T4.Prod[(T4.Prod$Dawn..Dusk == "dawn"),] # dawn-1 measurements
T4.Dusk<-T4.Prod[(T4.Prod$Dawn..Dusk == "dusk"),] # dusk measurements

T4.Dawn2<-YSI.T4[(YSI.T4$Date == "2022-02-03"),] # dawn-2 measurements, following AM

# make new dataframe
T4.O2<-(T4.Dawn1[,c(2,4:6)]) 
T4.O2$dawn1<-T4.Dawn1$DO.percent
T4.O2$dusk1<-T4.Dusk$DO.percent
T4.O2$dawn2<-T4.Dawn2$DO.percent

T4.O2<- mutate(T4.O2, 
                NEP=dusk1 - dawn1,
                NER=dawn2 - dusk1) 

#sort
T4.O2<-T4.O2 %>% 
  arrange(Treatment, plant.mass..g) 

################ ################ ################
# combine T1  T2  T3 T4 timepoints
################ ################ ################
O2.tanks<-rbind(T1.O2,T2.O2, T3.O2, T4.O2)
```
  
TIME POINT 1: Change in O2% and NEP and NER --   
make figures of the data for timepoint 1  
```{r}
########## 
Fig.formatting<-(theme_classic()) +
  theme(text=element_text(size=10),
        axis.line=element_blank(),
        legend.text.align = 0,
        legend.text=element_text(size=10),
        #legend.title = element_blank(),
        panel.border = element_rect(fill=NA, colour = "black", size=1),
        aspect.ratio=1, 
        axis.ticks.length=unit(0.25, "cm"),
        axis.text.y=element_text(
          margin=unit(c(0.5, 0.5, 0.5, 0.5), "cm"), colour="black", size=10), 
        axis.text.x=element_text(
          margin=unit(c(0.5, 0.5, 0.5, 0.5), "cm"), colour="black", size=8)) +
  theme(legend.key.size = unit(0.4, "cm")) +
  theme(aspect.ratio=1.3) +
  theme(panel.spacing=unit(c(0, 0, 0, 0), "cm"))

#########################################################
############################################################################
###############################################################################################
###################

# total oxygen % plot for the 3 time points (dawn-dusk-dawn)

T1.AM1<-ggplot(T1.O2, aes(x=plant.mass..g, y=dawn1, color=Treatment)) + 
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  geom_hline(yintercept=0, linetype="longdash", color = "gray") +
  ggtitle("T1.dawn1")+
  coord_cartesian(ylim=c(-30, 150)) +
  ylab(expression(paste("O"[2],"%"))) +
  xlab("plant material (g)") +
  Fig.formatting

T1.PM1<-ggplot(T1.O2, aes(x=plant.mass..g, y=dusk1, color=Treatment)) + 
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  geom_hline(yintercept=0, linetype="longdash", color = "gray") +
  ggtitle("T1.dusk1")+
  coord_cartesian(ylim=c(-30, 150)) +
  ylab(expression(paste("O"[2],"%"))) +
  xlab("plant material (g)") +
  Fig.formatting

T1.AM2<-ggplot(T1.O2, aes(x=plant.mass..g, y=dawn2, color=Treatment)) + 
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  geom_hline(yintercept=0, linetype="longdash", color = "gray") +
  ggtitle("T1.dawn2")+
  coord_cartesian(ylim=c(-30, 150)) +
  ylab(expression(paste("O"[2],"%"))) +
  xlab("plant material (g)") +
  Fig.formatting

extract.legend <- get_legend(
  # create some space to the left of the legend
  T1.AM1 + theme(legend.box.margin = margin(0, 0, 0, 10)))

O2.T1<-plot_grid(
  T1.AM1+ theme(legend.position = "none"), 
  T1.PM1+ theme(legend.position = "none"), 
  T1.AM2+ theme(legend.position = "none"),
  extract.legend, 
  rel_widths = c(8,8,8,3), ncol=4)

ggsave("figures/O2.T1.pdf", height=5, width=9)

#################################
############################################
# net ecosystem production and respiration

#Time 1 NEP
NEP.T1<-ggplot(T1.O2, aes(x=plant.mass..g, y=NEP, color=Treatment)) + 
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  geom_hline(yintercept=0, linetype="longdash", color = "gray") +
  ylab(expression(paste("Net Production (", Delta, "O"[2],"%)"))) +
  coord_cartesian(ylim=c(-20, 50)) +
  xlab("plant material (g)")+
  Fig.formatting

#Time 1 NER
NER.T1<-ggplot(T1.O2, aes(x=plant.mass..g, y=NER, color=Treatment)) + 
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  geom_hline(yintercept=0, linetype="longdash", color = "gray") +
  coord_cartesian(ylim=c(-40, 10)) +
  ylab(expression(paste("Net Respiration (", Delta, "O"[2],"%)"))) +
  xlab("plant material (g)") +
  Fig.formatting


O2.change.T1<-plot_grid(
  NEP.T1+ theme(legend.position = "none") + ggtitle("Time 1"), 
  NER.T1+ theme(legend.position = "none") + ggtitle(""),
  extract.legend, 
  rel_widths = c(8,8,3), ncol=3)

ggsave("figures/O2.change.T1.pdf", height=5, width=8)
```
  
TIME POINT 2: Change in O2% and NEP and NER --
make figures of the data for timepoint 2
```{r}
############################################################
##############################################################################
#############################################################################################
###################

# total oxygen % plot for the 3 time points (dawn-dusk-dawn)

T2.AM1<-ggplot(T2.O2, aes(x=plant.mass..g, y=dawn1, color=Treatment)) + 
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  geom_hline(yintercept=0, linetype="longdash", color = "gray") +
  ggtitle("T2.dawn1")+
  coord_cartesian(ylim=c(-30, 150)) +
  ylab(expression(paste("O"[2],"%"))) +
  xlab("plant material (g)") +
  Fig.formatting

T2.PM1<-ggplot(T2.O2, aes(x=plant.mass..g, y=dusk1, color=Treatment)) + 
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  geom_hline(yintercept=0, linetype="longdash", color = "gray") +
  ggtitle("T2.dusk1")+
  coord_cartesian(ylim=c(-30, 150)) +
  ylab(expression(paste("O"[2],"%"))) +
  xlab("plant material (g)") +
  Fig.formatting

T2.AM2<-ggplot(T2.O2, aes(x=plant.mass..g, y=dawn2, color=Treatment)) + 
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  geom_hline(yintercept=0, linetype="longdash", color = "gray") +
  ggtitle("T2.dawn2")+
  coord_cartesian(ylim=c(-30, 150)) +
  ylab(expression(paste("O"[2],"%"))) +
  xlab("plant material (g)") +
  Fig.formatting

O2.T2<-plot_grid(
  T2.AM1+ theme(legend.position = "none"), 
  T2.PM1+ theme(legend.position = "none"),
  T2.AM2+ theme(legend.position = "none"),
  extract.legend, 
  rel_widths = c(8,8,8,3), ncol=4)

ggsave("figures/O2.T2.pdf", height=5, width=9)


############################################################
##############################################################################################
# net ecosystem production and respiration

#Time 2 NEP
NEP.T2<-ggplot(T2.O2, aes(x=plant.mass..g, y=NEP, color=Treatment)) + 
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  geom_hline(yintercept=0, linetype="longdash", color = "gray") +
  coord_cartesian(ylim=c(-20, 50)) +
  ylab(expression(paste("Net Production (", Delta, "O"[2],"%)"))) +
  xlab("plant material (g)") +
  Fig.formatting

#Time 2 NER
NER.T2<-ggplot(T2.O2, aes(x=plant.mass..g, y=NER, color=Treatment)) + 
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment),alpha=0.1) +
  geom_hline(yintercept=0, linetype="longdash", color = "gray") +
  coord_cartesian(ylim=c(-40, 10)) +
  ylab(expression(paste("Net Respiration (", Delta, "O"[2],"%)"))) +
  xlab("plant material (g)") +
  Fig.formatting

O2.change.T2<-plot_grid(
  NEP.T2+ theme(legend.position = "none") + ggtitle("Time 2"),
  NER.T2+ theme(legend.position = "none") + ggtitle(""),
  extract.legend, 
  rel_widths = c(8,8,3), ncol=3)

ggsave("figures/O2.change.T2.pdf", height=5, width=8)


```

TIME POINT 3: Change in O2% and NEP and NER --
make figures of the data for timepoint 3
```{r}
############################################################
##############################################################################
#############################################################################################
###################

# total oxygen % plot for the 3 time points (dawn-dusk-dawn)

T3.AM1<-ggplot(T3.O2, aes(x=plant.mass..g, y=dawn1, color=Treatment)) + 
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  geom_hline(yintercept=0, linetype="longdash", color = "gray") +
  ggtitle("T3.dawn1")+
  coord_cartesian(ylim=c(-30, 150)) +
  ylab(expression(paste("O"[2],"%"))) +
  xlab("plant material (g)") +
  Fig.formatting

T3.PM1<-ggplot(T3.O2, aes(x=plant.mass..g, y=dusk1, color=Treatment)) + 
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  geom_hline(yintercept=0, linetype="longdash", color = "gray") +
  ggtitle("T3.dusk1")+
  coord_cartesian(ylim=c(-30, 150)) +
  ylab(expression(paste("O"[2],"%"))) +
  xlab("plant material (g)") +
  Fig.formatting

T3.AM2<-ggplot(T3.O2, aes(x=plant.mass..g, y=dawn2, color=Treatment)) + 
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  geom_hline(yintercept=0, linetype="longdash", color = "gray") +
  ggtitle("T3.dawn2")+
  coord_cartesian(ylim=c(-30, 150)) +
  ylab(expression(paste("O"[2],"%"))) +
  xlab("plant material (g)") +
  Fig.formatting

O2.T3<-plot_grid(
  T3.AM1+ theme(legend.position = "none"), 
  T3.PM1+ theme(legend.position = "none"),
  T3.AM2+ theme(legend.position = "none"),
  extract.legend, 
  rel_widths = c(8,8,8,3), ncol=4)

ggsave("figures/O2.T3.pdf", height=5, width=9)


############################################################
##############################################################################################
# net ecosystem production and respiration

#Time 3 NEP
NEP.T3<-ggplot(T3.O2, aes(x=plant.mass..g, y=NEP, color=Treatment)) + 
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  geom_hline(yintercept=0, linetype="longdash", color = "gray") +
  coord_cartesian(ylim=c(-20, 50)) +
  ylab(expression(paste("Net Production (", Delta, "O"[2],"%)"))) +
  xlab("plant material (g)") +
  Fig.formatting

#Time 3 NER
NER.T3<-ggplot(T3.O2, aes(x=plant.mass..g, y=NER, color=Treatment)) + 
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment),alpha=0.1) +
  geom_hline(yintercept=0, linetype="longdash", color = "gray") +
  coord_cartesian(ylim=c(-40, 10)) +
  ylab(expression(paste("Net Respiration (", Delta, "O"[2],"%)"))) +
  xlab("plant material (g)") +
  Fig.formatting

O2.change.T3<-plot_grid(
  NEP.T3+ theme(legend.position = "none") + ggtitle("Time 3"),
  NER.T3+ theme(legend.position = "none") + ggtitle(""),
  extract.legend, 
  rel_widths = c(8,8,3), ncol=3)

ggsave("figures/O2.change.T3.pdf", height=5, width=8)


```
  
TIME POINT 4: Change in O2% and NEP and NER --
make figures of the data for timepoint 4
```{r}
############################################################
##############################################################################
#############################################################################################
###################

# total oxygen % plot for the 3 time points (dawn-dusk-dawn)

T4.AM1<-ggplot(T4.O2, aes(x=plant.mass..g, y=dawn1, color=Treatment)) + 
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  geom_hline(yintercept=0, linetype="longdash", color = "gray") +
  ggtitle("T4.dawn1")+
  coord_cartesian(ylim=c(-30, 150)) +
  ylab(expression(paste("O"[2],"%"))) +
  xlab("plant material (g)") +
  Fig.formatting

T4.PM1<-ggplot(T4.O2, aes(x=plant.mass..g, y=dusk1, color=Treatment)) + 
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  geom_hline(yintercept=0, linetype="longdash", color = "gray") +
  ggtitle("T4.dusk1")+
  coord_cartesian(ylim=c(-30, 150)) +
  ylab(expression(paste("O"[2],"%"))) +
  xlab("plant material (g)") +
  Fig.formatting

T4.AM2<-ggplot(T4.O2, aes(x=plant.mass..g, y=dawn2, color=Treatment)) + 
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  geom_hline(yintercept=0, linetype="longdash", color = "gray") +
  ggtitle("T4.dawn2")+
  coord_cartesian(ylim=c(-30, 150)) +
  ylab(expression(paste("O"[2],"%"))) +
  xlab("plant material (g)") +
  Fig.formatting

O2.T4<-plot_grid(
  T4.AM1+ theme(legend.position = "none"), 
  T4.PM1+ theme(legend.position = "none"),
  T4.AM2+ theme(legend.position = "none"),
  extract.legend, 
  rel_widths = c(8,8,8,3), ncol=4)

ggsave("figures/O2.T4.pdf", height=5, width=9)


############################################################
##############################################################################################
# net ecosystem production and respiration

#Time 4 NEP
NEP.T4<-ggplot(T4.O2, aes(x=plant.mass..g, y=NEP, color=Treatment)) + 
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  geom_hline(yintercept=0, linetype="longdash", color = "gray") +
  coord_cartesian(ylim=c(-20, 50)) +
  ylab(expression(paste("Net Production (", Delta, "O"[2],"%)"))) +
  xlab("plant material (g)") +
  Fig.formatting

#Time 4 NER
NER.T4<-ggplot(T4.O2, aes(x=plant.mass..g, y=NER, color=Treatment)) + 
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment),alpha=0.1) +
  geom_hline(yintercept=0, linetype="longdash", color = "gray") +
  coord_cartesian(ylim=c(-40, 10)) +
  ylab(expression(paste("Net Respiration (", Delta, "O"[2],"%)"))) +
  xlab("plant material (g)") +
  Fig.formatting

O2.change.T4<-plot_grid(
  NEP.T4+ theme(legend.position = "none") + ggtitle("Time 4"),
  NER.T4+ theme(legend.position = "none") + ggtitle(""),
  extract.legend, 
  rel_widths = c(8,8,3), ncol=3)

ggsave("figures/O2.change.T4.pdf", height=5, width=8)

########## ########## ########## ########## ########## 
########## compile all figs

##### All dawn-dusk-dawn measurements
O2change.alltimes<-plot_grid(
  O2.T1,
  O2.T2,
  O2.T3,
  O2.T4, ncol=1)

ggsave("figures/O2change.alltimes.pdf", height=12, width=10)


##### All photosynthesis and respiration

PnR.alltimes<-plot_grid(
  O2.change.T1,
  O2.change.T2,
  O2.change.T3,
  O2.change.T4, ncol=1)

ggsave("figures/PnR.alltimes.pdf", height=12, width=8)


```
  
  
###Chlorophyll  
import 
```{r}
## import treatment IDs
chl.dat<- read.csv("data/Pyro_chlorophyll.csv")  
chl.dat<-chl.dat[!(chl.dat$Time.point=="First.round"),] # remove first round from aborted mission


#plot

chla.T0<-ggplot(chl.dat[(chl.dat$Time.point=="T0"),], aes(x=plant.mass..g, y=chla..ug.L, color=Treatment)) +
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  coord_cartesian(ylim=c(0, 20)) +
  ggtitle("Time-0") +
  ylab("chlorophyll (ug/L)") +
  xlab("plant material (g)") +
  Fig.formatting


chla.T1<-ggplot(chl.dat[(chl.dat$Time.point=="T1"),], aes(x=plant.mass..g, y=chla..ug.L, color=Treatment)) +
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  coord_cartesian(ylim=c(0, 20)) +
  ggtitle("Time-1") +
  ylab("chlorophyll (ug/L)") +
  xlab("plant material (g)") +
  Fig.formatting

chla.T2<-ggplot(chl.dat[(chl.dat$Time.point=="T2"),], aes(x=plant.mass..g, y=chla..ug.L, color=Treatment)) +
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  coord_cartesian(ylim=c(0, 20)) +
  ggtitle("Time-2") +
  ylab("chlorophyll (ug/L)") +
  xlab("plant material (g)") +
  Fig.formatting

chla.T3<-ggplot(chl.dat[(chl.dat$Time.point=="T3"),], aes(x=plant.mass..g, y=chla..ug.L, color=Treatment)) +
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  coord_cartesian(ylim=c(0, 20)) +
  ggtitle("Time-3") +
  ylab("chlorophyll (ug/L)") +
  xlab("plant material (g)") +
  Fig.formatting

chla.T4<-ggplot(chl.dat[(chl.dat$Time.point=="T4"),], aes(x=plant.mass..g, y=chla..ug.L, color=Treatment)) +
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  coord_cartesian(ylim=c(0, 20)) +
  ggtitle("Time-4") +
  ylab("chlorophyll (ug/L)") +
  xlab("plant material (g)") +
  Fig.formatting


chl.plots<-plot_grid(
  chla.T0+ theme(legend.position = "none"),
  chla.T1+ theme(legend.position = "none"),
  chla.T2+ theme(legend.position = "none"),
  chla.T3+ theme(legend.position = "none"),
  chla.T4+ theme(legend.position = "none"),
  extract.legend, 
  rel_widths = c(8,8,8,8,8,3), ncol=6)


chl.plots

ggsave("figures/chla.plots.pdf", height=5, width=12)




#remove outlier
chl.out.rem<-chl.dat[!(chl.dat$chla..ug.L>=10),]


chla.T0<-ggplot(chl.out.rem[(chl.out.rem$Time.point=="T0"),], aes(x=plant.mass..g, y=chla..ug.L, color=Treatment)) +
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  coord_cartesian(ylim=c(0, 12)) +
  ggtitle("Time-0") +
  ylab("chlorophyll (ug/L)") +
  xlab("plant material (g)") +
  Fig.formatting


chla.T1<-ggplot(chl.out.rem[(chl.out.rem$Time.point=="T1"),], aes(x=plant.mass..g, y=chla..ug.L, color=Treatment)) +
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  coord_cartesian(ylim=c(0, 12)) +
  ggtitle("Time-1") +
  ylab("chlorophyll (ug/L)") +
  xlab("plant material (g)") +
  Fig.formatting

chla.T2<-ggplot(chl.out.rem[(chl.out.rem$Time.point=="T2"),], aes(x=plant.mass..g, y=chla..ug.L, color=Treatment)) +
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  coord_cartesian(ylim=c(0, 12)) +
  ggtitle("Time-2") +
  ylab("chlorophyll (ug/L)") +
  xlab("plant material (g)") +
  Fig.formatting

chla.T3<-ggplot(chl.out.rem[(chl.out.rem$Time.point=="T3"),], aes(x=plant.mass..g, y=chla..ug.L, color=Treatment)) +
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  coord_cartesian(ylim=c(0, 12)) +
  ggtitle("Time-3") +
  ylab("chlorophyll (ug/L)") +
  xlab("plant material (g)") +
  Fig.formatting

chla.T4<-ggplot(chl.out.rem[(chl.out.rem$Time.point=="T4"),], aes(x=plant.mass..g, y=chla..ug.L, color=Treatment)) +
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  coord_cartesian(ylim=c(0, 12)) +
  ggtitle("Time-4") +
  ylab("chlorophyll (ug/L)") +
  xlab("plant material (g)") +
  Fig.formatting


chl.plots<-plot_grid(
  chla.T0+ theme(legend.position = "none"),
  chla.T1+ theme(legend.position = "none"),
  chla.T2+ theme(legend.position = "none"),
  chla.T3+ theme(legend.position = "none"),
  chla.T4+ theme(legend.position = "none"),
  extract.legend, 
  rel_widths = c(8,8,8,8,8,3), ncol=6)


chl.plots

ggsave("figures/chla.plots2.pdf", height=5, width=12)



```
  
  
###DOC  
import and do a loop to clean up all files and make stacked data in single df
```{r}
## import treatment IDs
IDs<-read.csv("data/treatment.IDs.csv")

##### grab files in a list
Total.DOC.files <- list.files(path="data/DOC.TN", pattern = "csv$", full.names = T)
Total.DOC.files

##### what are the file names, sans extensions using package 'tools'
file.names<-file_path_sans_ext(list.files(path="data/DOC.TN", pattern = "csv$", full.names = F))

############ formatting all data in for loop
  for(i in 1:length(Total.DOC.files))
    {
  data<-read.csv(Total.DOC.files[i], sep=",")
  data<-data[,c(1,3,5)] # removed columns we don't need
  data$File<-Total.DOC.files[i]
  colnames(data)<-c("Tank", "NPOC..mg.L", "TN..mg.L", "File")
  data$Tank<-as.numeric(as.character(data$Tank)) # make the column of samples all numeric
  data <- data[!is.na(as.numeric(as.character(data$Tank))),] # remove all rows that aren't numeric/tanks
  data$Treatment<-IDs$Treatment
  data$plant.mass..g<-IDs$plant.mass..g
  make.names(assign(paste(file.names[i], sep=""), data)) # make the file name the name of new df for loop df
  }
########## this is the end of the loop


ls() #see all dfs you've made, the above will be df matching their file names

#Combine files from loop to single df
DOC.df<-rbind(DOC_T0_11052021, DOC_T1_11172021, DOC_T2_12082021, DOC_T3_01072022)

DOC.df$File<-substr(DOC.df$File, 13, 27) #give the 10th-24th character of the file name, removing the rest
#alternatively
# remove the 9 letters ('^.) at start 
# remove the 4 letters (.$') at end
#DOC.df$File<-gsub('^.........|....$', '', DOC.df$File) 

# if equals DOC_T0_11052021 then, T0, if not then T1
DOC.df$Time.point<- as.factor(ifelse(DOC.df$File=="DOC_T0_11052021", "T0", 
                              ifelse(DOC.df$File=="DOC_T1_11172021", "T1",
                                     ifelse(DOC.df$File=="DOC_T2_12082021", "T2", 
                                     "T3"))))
  
#rearrange
DOC.df<- DOC.df %>% 
  select(File, Time.point, Treatment, Tank, plant.mass..g, NPOC..mg.L, TN..mg.L) 

```

DOC plots
```{r}
NPOC.T0<-ggplot(DOC.df[(DOC.df$Time.point=="T0"),], aes(x=plant.mass..g, y=NPOC..mg.L, color=Treatment)) +
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  coord_cartesian(ylim=c(0, 60)) +
  ggtitle("Time-0") +
  ylab("NPOC (mg/L)") +
  xlab("plant material (g)") +
  Fig.formatting

NPOC.T1<-ggplot(DOC.df[(DOC.df$Time.point=="T1"),], aes(x=plant.mass..g, y=NPOC..mg.L, color=Treatment)) + 
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  coord_cartesian(ylim=c(0, 60)) +
  ggtitle("Time-1") +
  ylab("NPOC (mg/L)") +
  xlab("plant material (g)") +
  Fig.formatting
  
NPOC.T2<-ggplot(DOC.df[(DOC.df$Time.point=="T2"),], aes(x=plant.mass..g, y=NPOC..mg.L, color=Treatment)) + 
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  coord_cartesian(ylim=c(0, 60)) +
  ggtitle("Time-2") +
  ylab("NPOC (mg/L)") +
  xlab("plant material (g)") +
  Fig.formatting

  
NPOC.T3<-ggplot(DOC.df[(DOC.df$Time.point=="T3"),], aes(x=plant.mass..g, y=NPOC..mg.L, color=Treatment)) + 
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  coord_cartesian(ylim=c(0, 60)) +
  ggtitle("Time-3") +
  ylab("NPOC (mg/L)") +
  xlab("plant material (g)") +
  Fig.formatting

NPOC.plots<-plot_grid(
  NPOC.T0+ theme(legend.position = "none"),
  NPOC.T1+ theme(legend.position = "none"),
  NPOC.T2+ theme(legend.position = "none"),
  NPOC.T3+ theme(legend.position = "none"),
  extract.legend, 
  rel_widths = c(8,8,8,8,3), ncol=5)


NPOC.plots

ggsave("figures/NPOC.pdf", height=5, width=12)

```

Total N plots
```{r}
TN.T0<-ggplot(DOC.df[(DOC.df$Time.point=="T0"),], aes(x=plant.mass..g, y=TN..mg.L, color=Treatment)) + 
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  coord_cartesian(ylim=c(0, 2)) +
  ggtitle("Time-0") +
  ylab("TN (mg/L)") +
  xlab("plant material (g)") +
  Fig.formatting
 
TN.T1<-ggplot(DOC.df[(DOC.df$Time.point=="T1"),], aes(x=plant.mass..g, y=TN..mg.L, color=Treatment)) + 
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  coord_cartesian(ylim=c(0, 2)) +
  ggtitle("Time-1") +
  ylab("TN (mg/L)") +
  xlab("plant material (g)") +
  Fig.formatting

TN.T2<-ggplot(DOC.df[(DOC.df$Time.point=="T2"),], aes(x=plant.mass..g, y=TN..mg.L, color=Treatment)) + 
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  coord_cartesian(ylim=c(0, 2)) +
  ggtitle("Time-2") +
  ylab("TN (mg/L)") +
  xlab("plant material (g)") +
  Fig.formatting

TN.T3<-ggplot(DOC.df[(DOC.df$Time.point=="T3"),], aes(x=plant.mass..g, y=TN..mg.L, color=Treatment)) + 
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  coord_cartesian(ylim=c(0, 2)) +
  ggtitle("Time-3") +
  ylab("TN (mg/L)") +
  xlab("plant material (g)") +
  Fig.formatting
  
TN.plots<-plot_grid(
  TN.T0+ theme(legend.position = "none"),
  TN.T1+ theme(legend.position = "none"),
  TN.T2+ theme(legend.position = "none"),
  TN.T3+ theme(legend.position = "none"),
  extract.legend, 
  rel_widths = c(8,8,8,8,3), ncol=5)

ggsave("figures/TN.pdf", height=5, width=15)

```


### UV/BIO DOC

```{r}

##### grab files in a list
UVBio.files <- list.files(path="data/UVBIO", pattern = "csv$", full.names = T)
UVBio.files

##### what are the file names, sans extensions using package 'tools'
file.names<-file_path_sans_ext(list.files(path="data/UVBIO", pattern = "csv$", full.names = F))
file.names


############ formatting all data in for loop
  for(i in 1:length(UVBio.files))
    {
  data<-read.csv(UVBio.files[i], sep=",", skip=12) #skip 13 rows where standards are
  data<-data[,c(1,3,5,7, 9)] # removed columns we don't need
  data$File<-UVBio.files[i]
  colnames(data)<-c("Tank", "NPOC..mg.L", "TN..mg.L", "Start.NPOC", "Start.TN", "File")
  data<-subset(data, Tank!="DI" & Tank!="UP") # remove the standards within sample run
  data$Tank<-as.character(data$Tank)
  data$Treatment<-IDs$Treatment # add treatment info
  data$plant.mass..g<-IDs$plant.mass..g # add plant biomass info
  make.names(assign(paste(file.names[i], sep=""), data)) # make the file name the name of new df for loop df
  }
########## this is the end of the loop


ls() #see all dfs you've made, the above will be df matching their file names

 


#Combine files from loop to single df
UVBIO.df<-rbind(DOC_NoUV_T1_111721, DOC_UV_T1_111721, DOC_NoUV_T2_120821, DOC_UV_T2_120821)

# make new factors for filtered (F/NF) and UV (+/-)
# ** caution! Make sure this matches sheet layout after the loop!

UVBIO.df$Filter.Trt<-rep(c("Filtered", "Unfiltered"), each=30) # repeat 30 times
UVBIO.df$UV.Trt<-rep(c("NoUV", "UV"), each=60) # repeat 60 times

#fix tank ID name
UVBIO.df$Tank[UVBIO.df$Tank == "NO UV F1" | 
                UVBIO.df$Tank == "NO UV NF1" |
                UVBIO.df$Tank == "UV F1" |
                UVBIO.df$Tank == "UV NF1"] <- "1"

# extract out the file names
UVBIO.df$File<-gsub('^...........|....$', '', UVBIO.df$File) 

# add a time point
# if equals DOC_UV_T1_111721 then, T1, if not then "nothing"
UVBIO.df$Time.point<- as.factor(ifelse(UVBIO.df$File=="DOC_UV_T1_111721", "T1",
  ifelse(UVBIO.df$File=="DOC_NoUV_T1_111721", "T1",
          ifelse(UVBIO.df$File=="DOC_UV_T2_12082021", "T2",
                ifelse(UVBIO.df$File=="DOC_NoUV_T2_12082021.csv", "T2", "T2")))))
  
#rearrange
UVBIO.df<- UVBIO.df %>% 
  select(File, Time.point, Treatment, Filter.Trt, UV.Trt, Tank, plant.mass..g, NPOC..mg.L, TN..mg.L, Start.NPOC, Start.TN) 


# make levels for plotting and stats

```

plot UV Bio

```{r}
UVBIO.df$Burn.Filt.UV<-interaction(UVBIO.df$Treatment, UVBIO.df$Filter.Trt, UVBIO.df$UV.Trt)
UVBIO.df$Filt.UV<-interaction(UVBIO.df$Filter.Trt, UVBIO.df$UV.Trt)

# correct to make factor levels
UVBIO.df$Filter.Trt<-as.factor(UVBIO.df$Filter.Trt)
UVBIO.df$UV.Trt <-as.factor(UVBIO.df$UV.Trt)
UVBIO.df$Tank <-as.factor(UVBIO.df$Tank)

# make new variables
# (1) change in NPOC over the 7d incubation
UVBIO.df$change.NPOC<-UVBIO.df$Start.NPOC - UVBIO.df$NPOC..mg.L
UVBIO.df$change.TN<- UVBIO.df$Start.TN - UVBIO.df$TN..mg.L

# (1) % change in NPOC over the 7d incubation
UVBIO.df<-mutate (UVBIO.df, percent.change.NPOC=(change.NPOC/Start.NPOC)*100)
UVBIO.df<-mutate (UVBIO.df, percent.change.TN=(change.TN/Start.TN)*100)

# make some plotting datafames
# these are all separated by the 4 incubation treatment levels
Filt.UV <-UVBIO.df[UVBIO.df$Filter.Trt=="Filtered" & UVBIO.df$UV.Trt=="UV",]

Filt.NoUV<-UVBIO.df[(UVBIO.df$Filter.Trt=="Filtered" & UVBIO.df$UV.Trt=="NoUV"),]

NFilt.UV <-UVBIO.df[UVBIO.df$Filter.Trt=="Unfiltered" & UVBIO.df$UV.Trt=="UV",]

NFilt.NoUV <-UVBIO.df[UVBIO.df$Filter.Trt=="Unfiltered" & UVBIO.df$UV.Trt=="NoUV",]


```
  
- plots of 2: % change of DOC
```{r}
UV.TRT.plot<-ggplot(UVBIO.df, aes(x=plant.mass..g, y=percent.change.NPOC, color=Treatment)) + 
  geom_point(alpha=0.5) +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  scale_linetype_manual(values=c("solid", "twodash")) +
  geom_smooth(method=loess, aes(fill=Treatment, linetype=UV.Trt), alpha=0.1) +
  ggtitle("No-UV vs. UV") +
  coord_cartesian(ylim=c(-5, 40)) +
  ylab(expression(paste("% ", Delta, " NPOC mg/L"))) +
  xlab("plant material (g)") +
  Fig.formatting

FILTER.TRT.plot<-ggplot(UVBIO.df, aes(x=plant.mass..g, y=percent.change.NPOC, color=Treatment)) + 
  geom_point(alpha=0.5) +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  scale_linetype_manual(values=c("solid", "twodash")) +
  geom_smooth(method=loess, aes(fill=Treatment, linetype=Filter.Trt), alpha=0.1) +
  ggtitle("Filtered vs. Unfiltered") +
  coord_cartesian(ylim=c(-5, 40)) +
  ylab(expression(paste("% ", Delta, " NPOC mg/L"))) +
  xlab("plant material (g)") +
  Fig.formatting


##### combine
UVBio.plots2<-plot_grid(
  UV.TRT.plot,
  FILTER.TRT.plot,  ncol=2)

ggsave("figures/UVBio.plots.percent2.pdf", height=5, width=9)

```

- another take on the UV-BIO separated by burning treatment
```{r}
UV.TRT.B.plot<-ggplot(UVBIO.df[(UVBIO.df$Treatment=="burned"),], 
                       aes(x=plant.mass..g, y=percent.change.NPOC, color=Filt.UV)) + 
  geom_point(alpha=0.5) +
  scale_color_manual(values = c("coral", "yellowgreen", "peru", "seagreen")) + 
  scale_fill_manual(values = c("coral", "yellowgreen", "peru", "seagreen")) +
  geom_smooth(method=loess, aes(fill=Filt.UV), alpha=0.1) +
  ggtitle("Unburned") +
  coord_cartesian(ylim=c(-5, 40)) +
  ylab(expression(paste("% ", Delta, " NPOC mg/L"))) +
  xlab("plant material (g)") +
  Fig.formatting

UV.TRT.UB.plot<-ggplot(UVBIO.df[(UVBIO.df$Treatment=="unburned"),], 
                       aes(x=plant.mass..g, y=percent.change.NPOC, color=Filt.UV)) + 
  geom_point(alpha=0.5) +
  scale_color_manual(values = c("coral", "yellowgreen", "peru", "seagreen")) + 
  scale_fill_manual(values = c("coral", "yellowgreen", "peru", "seagreen")) +
  geom_smooth(method=loess, aes(fill=Filt.UV), alpha=0.1) +
  ggtitle("Burned") +
  coord_cartesian(ylim=c(-5, 40)) +
  ylab(expression(paste("% ", Delta, " NPOC mg/L"))) +
  xlab("plant material (g)") +
  Fig.formatting


extract.legend2 <- get_legend(
  # create some space to the left of the legend
  UV.TRT.UB.plot + theme(legend.box.margin = margin(0, 0, 0, 10)))

NFilt.NoUV <-UVBIO.df[UVBIO.df$Filter.Trt=="Unfiltered" & UVBIO.df$UV.Trt=="NoUV",]


##### combine
UVBio.plots3<-plot_grid(
  UV.TRT.B.plot + theme(legend.position = "none"),
  UV.TRT.UB.plot + theme(legend.position = "none"),
  extract.legend2,
  rel_widths = c(8,8,3), ncol=3)



ggsave("figures/UVBio.plots.percent3.pdf", height=5, width=9)


```

<<<<<<< HEAD
 
=======
- plot of 4: UV-BIO as % change in DOC
```{r}
################## plots for total DOC concentration over incubation period

# filtered UV plot
Filt.UV.fig<-ggplot(Filt.UV, aes(x=plant.mass..g, y=NPOC..mg.L, color=Treatment)) + 
  geom_point() +
   facet_grid(.~ Time.point) +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  ggtitle("Filt:UV") +
  ylab(expression(paste(" NPOC mg/L"))) +
  xlab("plant material (g)") 

# filtered NoUV plot
F.NoUV.fig<-ggplot(Filt.NoUV, aes(x=plant.mass..g, y=NPOC..mg.L, color=Treatment)) + 
  geom_point() +
   facet_grid(.~ Time.point) +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  ggtitle("Filt:NoUV") +
  ylab(expression(paste("NPOC mg/L"))) +
  xlab("plant material (g)") 


# nofiltered UV plot
NFilt.UV.fig<-ggplot(NFilt.UV, aes(x=plant.mass..g, y=NPOC..mg.L, color=Treatment)) + 
  geom_point() +
   facet_grid(.~ Time.point) +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  ggtitle("NoFilt:UV") +
  ylab(expression(paste("NPOC mg/L"))) +
  xlab("plant material (g)") 


# nofiltered NoUV plot
NFilt.NoUV.fig<-ggplot(NFilt.NoUV, aes(x=plant.mass..g, y=NPOC..mg.L, color=Treatment)) +
  geom_point() +
  facet_wrap(.~ Time.point) +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  ggtitle("NoFilt:NoUV") +
  ylab(expression(paste("NPOC mg/L"))) +
  xlab("plant material (g)") 

##### combine
UVBio.DOC.timeseries<-plot_grid(
  Filt.UV.fig + theme(legend.position = "none"),
  F.NoUV.fig + theme(legend.position = "none"),
  NFilt.UV.fig + theme(legend.position = "none"),
  NFilt.NoUV.fig + theme(legend.position = "none"),
  extract.legend, 
  rel_widths = c(8,8,8,8,3), ncol=5)

UVBio.DOC.timeseries

ggsave("figures/UVBio.DOC.timeseries.pdf", height=4, width=15)

################## plots for change in DOC over incubation period

# filtered UV plot
change.Filt.UV.fig<-ggplot(Filt.UV, aes(x=plant.mass..g, y=change.NPOC, color=Treatment)) + 
  geom_point() +
   facet_grid(.~ Time.point) +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  ggtitle("Filt:UV") +
  ylab(expression(paste(Delta, "NPOC mg/L"))) +
  xlab("plant material (g)") 

# filtered NoUV plot
change.F.NoUV.fig<-ggplot(Filt.NoUV, aes(x=plant.mass..g, y=change.NPOC, color=Treatment)) + 
  geom_point() +
   facet_grid(.~ Time.point) +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  ggtitle("Filt:NoUV") +
  ylab(expression(paste(Delta, "NPOC mg/L"))) +
  xlab("plant material (g)") 


# nofiltered UV plot
change.NFilt.UV.fig<-ggplot(NFilt.UV, aes(x=plant.mass..g, y=change.NPOC, color=Treatment)) + 
  geom_point() +
   facet_grid(.~ Time.point) +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  ggtitle("NoFilt:UV") +
  ylab(expression(paste(Delta, "NPOC mg/L"))) +
  xlab("plant material (g)") 


# nofiltered NoUV plot
change.NFilt.NoUV.fig<-ggplot(NFilt.NoUV, aes(x=plant.mass..g, y=change.NPOC, color=Treatment)) +
  geom_point() +
  facet_wrap(.~ Time.point) +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  ggtitle("NoFilt:NoUV") +
  ylab(expression(paste(Delta, "NPOC mg/L"))) +
  xlab("plant material (g)") 

##### combine
change.UVBio.DOC<-plot_grid(
  change.Filt.UV.fig + theme(legend.position = "none"),
  change.F.NoUV.fig + theme(legend.position = "none"),
  change.NFilt.UV.fig + theme(legend.position = "none"),
  change.NFilt.NoUV.fig + theme(legend.position = "none"),
  extract.legend, 
  rel_widths = c(8,8,8,8,3), ncol=5)

change.UVBio.DOC
ggsave("figures/change.UVBio.DOC.pdf", height=5, width=15)

################## plots for % change in DOC over incubation period

# filtered UV plot
percent.change.Filt.UV.fig<-ggplot(Filt.UV, aes(x=plant.mass..g, y=percent.change.NPOC, color=Treatment)) + 
  geom_point() +
   facet_grid(.~ Time.point) +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  ggtitle("Filt:UV") +
  ylab(expression(paste("%", Delta, "NPOC mg/L"))) +
  xlab("plant material (g)") 

# filtered NoUV plot
percent.change.F.NoUV.fig<-ggplot(Filt.NoUV, aes(x=plant.mass..g, y=percent.change.NPOC, color=Treatment)) + 
  geom_point() +
   facet_grid(.~ Time.point) +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  ggtitle("Filt:NoUV") +
  ylab(expression(paste("%", Delta, "NPOC mg/L"))) +
  xlab("plant material (g)") 


# nofiltered UV plot
percent.change.NFilt.UV.fig<-ggplot(NFilt.UV, aes(x=plant.mass..g, y=percent.change.NPOC, color=Treatment)) + 
  geom_point() +
   facet_grid(.~ Time.point) +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  ggtitle("NoFilt:UV") +
  ylab(expression(paste("%", Delta, "NPOC mg/L"))) +
  xlab("plant material (g)") 


# nofiltered NoUV plot
percent.change.NFilt.NoUV.fig<-ggplot(NFilt.NoUV, aes(x=plant.mass..g, y=percent.change.NPOC, color=Treatment)) +
  geom_point() +
  facet_wrap(.~ Time.point) +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  ggtitle("NoFilt:NoUV") +
  ylab(expression(paste("%", Delta, "NPOC mg/L"))) +
  xlab("plant material (g)") 

##### combine
percent.change.UVBio.DOC<-plot_grid(
  percent.change.Filt.UV.fig + theme(legend.position = "none"),
  percent.change.F.NoUV.fig + theme(legend.position = "none"),
  percent.change.NFilt.UV.fig + theme(legend.position = "none"),
  percent.change.NFilt.NoUV.fig + theme(legend.position = "none"),
  extract.legend, 
  rel_widths = c(8,8,8,8,3), ncol=5)

percent.change.UVBio.DOC
ggsave("figures/percent.change.UVBio.DOC.pdf", height=4, width=15)

```


change in TN over incubation period
```{r}

################## plots for TN concentration over incubation period

# filtered UV plot
TN.Filt.UV.fig<-ggplot(Filt.UV, aes(x=plant.mass..g, y=TN..mg.L, color=Treatment)) + 
  geom_point() +
   facet_grid(.~ Time.point) +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  ggtitle("Filt:UV") +
  ylab(expression(paste("TN mg/L"))) +
  xlab("plant material (g)") 

# filtered NoUV plot
TN.F.NoUV.fig<-ggplot(Filt.NoUV, aes(x=plant.mass..g, y=TN..mg.L, color=Treatment)) + 
  geom_point() +
   facet_grid(.~ Time.point) +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  ggtitle("Filt:NoUV") +
  ylab(expression(paste("TN mg/L"))) +
  xlab("plant material (g)") 


# nofiltered UV plot
TN.NFilt.UV.fig<-ggplot(NFilt.UV, aes(x=plant.mass..g, y=TN..mg.L, color=Treatment)) + 
  geom_point() +
   facet_grid(.~ Time.point) +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  ggtitle("NoFilt:UV") +
  ylab(expression(paste("TN mg/L"))) +
  xlab("plant material (g)") 


# nofiltered NoUV plot
TN.NFilt.NoUV.fig<-ggplot(NFilt.NoUV, aes(x=plant.mass..g, y=TN..mg.L, color=Treatment)) +
  geom_point() +
  facet_wrap(.~ Time.point) +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  ggtitle("NoFilt:NoUV") +
  ylab(expression(paste("TN mg/L"))) +
  xlab("plant material (g)") 

##### combine
TN.UVBio<-plot_grid(
  TN.Filt.UV.fig + theme(legend.position = "none"),
  TN.F.NoUV.fig + theme(legend.position = "none"),
  TN.NFilt.UV.fig + theme(legend.position = "none"),
  TN.NFilt.NoUV.fig + theme(legend.position = "none"),
  extract.legend, 
  rel_widths = c(8,8,8,8,3), ncol=5)

TN.UVBio
ggsave("figures/TN.UVBio.pdf", height=4, width=15)

################## plots for change in TN over incubation period

# filtered UV plot
change.TN.Filt.UV.fig<-ggplot(Filt.UV, aes(x=plant.mass..g, y=change.TN, color=Treatment)) + 
  geom_point() +
   facet_grid(.~ Time.point) +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  ggtitle("Filt:UV") +
  ylab(expression(paste(Delta, "TN mg/L"))) +
  xlab("plant material (g)") 

# filtered NoUV plot
change.TN.F.NoUV.fig<-ggplot(Filt.NoUV, aes(x=plant.mass..g, y=change.TN, color=Treatment)) + 
  geom_point() +
   facet_grid(.~ Time.point) +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  ggtitle("Filt:NoUV") +
  ylab(expression(paste(Delta, "TN mg/L"))) +
  xlab("plant material (g)") 


# nofiltered UV plot
change.TN.NFilt.UV.fig<-ggplot(NFilt.UV, aes(x=plant.mass..g, y=change.TN, color=Treatment)) + 
  geom_point() +
   facet_grid(.~ Time.point) +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  ggtitle("NoFilt:UV") +
  ylab(expression(paste(Delta, "TN mg/L"))) +
  xlab("plant material (g)") 


# nofiltered NoUV plot
change.TN.NFilt.NoUV.fig<-ggplot(NFilt.NoUV, aes(x=plant.mass..g, y=change.TN, color=Treatment)) +
  geom_point() +
  facet_wrap(.~ Time.point) +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  ggtitle("NoFilt:NoUV") +
  ylab(expression(paste(Delta, "TN mg/L"))) +
  xlab("plant material (g)") 

##### combine
change.TN.UVBio<-plot_grid(
  change.TN.Filt.UV.fig + theme(legend.position = "none"),
  change.TN.F.NoUV.fig + theme(legend.position = "none"),
  change.TN.NFilt.UV.fig + theme(legend.position = "none"),
  change.TN.NFilt.NoUV.fig + theme(legend.position = "none"),
  extract.legend, 
  rel_widths = c(8,8,8,8,3), ncol=5)

change.TN.UVBio
ggsave("figures/change.TN.UVBio.pdf", height=4, width=15)

################## plots for % change in TN over incubation period

# filtered UV plot
percent.change.TN.Filt.UV.fig<-ggplot(Filt.UV, aes(x=plant.mass..g, y=percent.change.TN, color=Treatment)) + 
  geom_point() +
   facet_grid(.~ Time.point) +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  ggtitle("Filt:UV") +
  ylab(expression(paste("%", Delta, "TN mg/L"))) +
  xlab("plant material (g)") 

# filtered NoUV plot
percent.change.TN.F.NoUV.fig<-ggplot(Filt.NoUV, aes(x=plant.mass..g, y=percent.change.TN, color=Treatment)) + 
  geom_point() +
   facet_grid(.~ Time.point) +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  ggtitle("Filt:NoUV") +
  ylab(expression(paste("%", Delta, "TN mg/L"))) +
  xlab("plant material (g)") 


# nofiltered UV plot
percent.change.TN.NFilt.UV.fig<-ggplot(NFilt.UV, aes(x=plant.mass..g, y=percent.change.TN, color=Treatment)) + 
  geom_point() +
   facet_grid(.~ Time.point) +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  ggtitle("NoFilt:UV") +
  ylab(expression(paste("%", Delta, "TN mg/L"))) +
  xlab("plant material (g)") 


# nofiltered NoUV plot
percent.change.TN.NFilt.NoUV.fig<-ggplot(NFilt.NoUV, aes(x=plant.mass..g, y=percent.change.TN, color=Treatment)) +
  geom_point() +
  facet_wrap(.~ Time.point) +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  ggtitle("NoFilt:NoUV") +
  ylab(expression(paste("%", Delta, "TN mg/L"))) +
  xlab("plant material (g)") 

##### combine
percent.change.TN.UVBio<-plot_grid(
  percent.change.TN.Filt.UV.fig + theme(legend.position = "none"),
  percent.change.TN.F.NoUV.fig + theme(legend.position = "none"),
  percent.change.TN.NFilt.UV.fig + theme(legend.position = "none"),
  percent.change.TN.NFilt.NoUV.fig + theme(legend.position = "none"),
  extract.legend, 
  rel_widths = c(8,8,8,8,3), ncol=5)

percent.change.TN.UVBio
ggsave("figures/percent.change.TN.UVBio.pdf", height=4, width=15)


```


This chunk looks at different statistics:

```{r}

##### Looking at the relationship between DOC types (D or UD)
# Read the file
Dilute<-read.csv("data/DOC.Dilutions/UndilutedvsDiluted.csv")
              
#Graph for SD and CV vs. biomass
SD_dilute<- ggplot(Dilute, aes(x=plant.mass..g, y=STDEV, color=Treatment)) + 
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  coord_cartesian(ylim=c(0, 2)) +
  ggtitle("STDEV between samples") +
  ylab("STDEV") +
  xlab("plant material (g)") +
  Fig.formatting

CV_dilute<- ggplot(Dilute, aes(x=plant.mass..g, y=CV, color=Treatment)) + 
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  coord_cartesian(ylim=c(0, 2)) +
  ggtitle("CV between samples") +
  ylab("CV") +
  xlab("plant material (g)") +
  Fig.formatting

Dilute_plots <- plot_grid(
  SD_dilute + theme(legend.position = "none"),
  CV_dilute + theme(legend.position = "none"),
  extract.legend, 
  rel_widths = c(9,9,3), ncol=3)
Dilute_plots

ggsave("figures/Dilute.plots.pdf", height=4, width=8)


```



```{r}

# nofiltered UV plot
ggplot(NFilt.UV, aes(x=plant.mass..g, y=change.NPOC, color=Treatment)) + 
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  coord_cartesian(ylim=c(-5, 40)) +
  ggtitle("T1: NoFilt:UV") +
  ylab(expression(paste("% ", Delta, " NPOC mg/L"))) +
  xlab("plant material (g)") +
  Fig.formatting

lm_y <- lm(change.NPOC~ plant.mass..g, data = NFilt.UV)
summary(lm_y)

plot(lm_y, which = 1)

# nofiltered NoUV plot
ggplot(NFilt.NoUV, aes(x=plant.mass..g, y=percent.change.NPOC, color=Treatment)) + 
  geom_point() +
  scale_color_manual(values = c("brown1", "mediumseagreen")) + 
  geom_smooth(method=loess, aes(fill=Treatment), alpha=0.1) +
  coord_cartesian(ylim=c(-5, 40)) +
  ggtitle("T1: NoFilt:NoUV") +
  ylab(expression(paste("% ", Delta, " NPOC mg/L"))) +
  xlab("plant material (g)") +
  Fig.formatting



```


